---
title: "Analysis of TCR repertoires of LckWT, LckCA and LckCA/KR mice"
author: "Veronika Niederlova"
date: "8/23/2022"
output: rmdformats::material
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE, error = TRUE, cache = TRUE)

library(tidyverse)
library(readr)
library(readxl)
library(cowplot)
library(patchwork)
library(immunarch)
library(pheatmap)
library(factoextra)
library(kableExtra)
library(SmartEDA)

ggtheme <- function() {
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10),
    text = element_text(size = 10, colour = "black"),
    legend.text = element_text(size = 10),
    legend.key.size =  unit(10, units = "points"),
    axis.text.x = element_text(angle = 90)) 
  }

binary <- function(x){
  x <- ifelse(x>0,1,0)
  return(x)
}

```


# Initial processing

This document describes the analysis of TCR repertoires for the manuscript *"Unique roles of coreceptor-bound LCK in helper and cytotoxic T cells"* by Horkova et al. The analysis workflow included: 

* sorting of CD4 and CD8 cells from lymph nodes and thymi of *Lck^WT^*, *Lck^CA^* and *Lck^CA/KR^* mice,
* isolation of RNA using [RNA Clean & Concentrator-5 kit](https://www.integratedsci.com.au/product/rna-clean-and-concentrator-5-kit.html)
* preparation of TCR-enriched libraries using the [NEBNext® Mouse Immune Sequencing Kit](https://international.neb.com/products/e6330-nebnext-immune-sequencing-kit-mouse#Product%20Information_Reagents-Supplied)
* sequencing the libraries on Illumina Miseq, 300bp paired-end reads
* demultiplexing reads in Illumina BaseSpace 
* processing fastq files using the recommended [pRESTO workflow](https://international.neb.com/-/media/nebus/files/nebnext-chart-pdfs/nebnext-immune-sequencing-data-analysis-workflow-mouse.pdf?rev=7f21b23a3ce444e3bc4dcd57282402ba&hash=C0AE101CBA6C31B2A897A207D9EA6A4D) on Galaxy with default parameters
* aligning processed fastq files using [MiXCR](https://mixcr.readthedocs.io/en/develop/)

Raw data are deposited in the Sequence Read Archive (PRJNA872031). Processed files can be downloaded [download from Zenodo]( ############# ). The Zenodo archive contains the following files, which are needed to run this script:

* merged outputs from MiXCR `merged_TCR_repertoires.csv`
* metadata file `metadata_Lck.csv`
* count tables with raw counts of CDR3 TRA and TRB sequences in each sample `count_table_tra.csv`, `"count_table_trb.csv"`
* TRA and TRB repertoires prepared for processing with the Immunarch package `immdata_tra.rds`, `immdata_trb.rds`

Let's read the required files:

```{r}
#
merged_repertoires <- read.csv("merged_TCR_repertoires.csv")
md <- read.csv("metadata_Lck.csv")

counts_tra <- read_csv("count_table_tra.csv")
counts_tra$...1 <- NULL

counts_trb <- read_csv("count_table_trb.csv")
counts_trb$...1 <- NULL


```

# Count tables

We will start with processing of the file `merged_TCR_repertoires.csv`. This file contains outputs of MiXCR software, i.e., all the assembled CDR3 clonotypes, their counts in each sample, nucleotide and amino-acid sequences, all V, D and J alignments with scores, together with some meta-data. the column `num_id` contains info about the sample number, which we will frequently use to attach meta-data to sample numbers.

Here, we summarize the variables in the file `merged_TCR_repertoires.csv`:

```{r}
kable(ExpData(data=merged_repertoires, type=2)) %>%
  kable_styling(full_width = F, font_size = 11, 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

First, we will generate summary count tables for TRA and TRB CDR3 sequences (these tables are provided as Supplementary table S1 in the manuscript). We will reorder the file so that there will be one column for each sample, and we will change the `num_id` of the samples to meaningful names. 

```{r}
count_table_tra <- merged_repertoires %>% 
  dplyr::filter(chain == "TRA") %>% 
  dplyr::select(aaSeqCDR3, num_id, cloneCount)

count_table_tra2 <- count_table_tra %>% 
  pivot_wider(names_from = num_id, values_from = cloneCount, values_fill = 0, values_fn = sum)

md2_tra <- merged_repertoires %>% filter(chain == "TRA") %>% group_by(aaSeqCDR3) %>% slice_head(n = 1)

excel_count_table_tra <- counts_tra %>% left_join(md2_tra %>% select(allVHitsWithScore, allDHitsWithScore, allJHitsWithScore))

md_to_join <- md %>% mutate(new_name = paste(Cell_type, Organ, Mouse_strain, paste0("Exp0",Exp)))  %>% select(num_id, new_name) %>% arrange(num_id)

order <- match(as.numeric(colnames(counts_tra)[2:29]), md_to_join$num_id)

colnames(excel_count_table_tra)[2:29] <- pull(md_to_join, new_name)[order]
excel_count_table_tra2 <- excel_count_table_tra %>% dplyr::select(1, 31, 32, 30, 22, 12, 25, 11, 6, 26, 13, 18, 4, 28, 17, 14, 8, 23, 9, 7, 16, 24, 2, 21, 5, 29, 10, 20, 19, 3, 15, 27)
```

Here, we show the resulting table for TRA:

```{r}
kable(excel_count_table_tra2 %>% head) %>%
  kable_styling(full_width = F, font_size = 11, 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


```{r}
count_table_trb <- merged_repertoires %>% 
  dplyr::filter(chain == "TRB") %>% 
  dplyr::select(aaSeqCDR3, num_id, cloneCount)

count_table_trb2 <- count_table_trb %>% 
  pivot_wider(names_from = num_id, values_from = cloneCount, values_fill = 0, values_fn = sum)

md2_trb <- merged_repertoires %>% 
  filter(chain == "TRB") %>% 
  group_by(aaSeqCDR3) %>% 
  slice_head(n = 1)

excel_count_table_trb <- counts_trb %>% 
  left_join(md2_trb %>% 
              select(allVHitsWithScore, allDHitsWithScore, allJHitsWithScore))

md_to_join <- md %>% 
  mutate(new_name = paste(Cell_type, Organ, Mouse_strain, paste0("Exp0",Exp))) %>% 
  select(num_id, new_name) %>% 
  arrange(num_id)

order <- match(as.numeric(colnames(counts_trb)[2:29]), md_to_join$num_id)

colnames(excel_count_table_trb)[2:29] <- pull(md_to_join, new_name)[order]

excel_count_table_trb2 <- excel_count_table_trb %>% 
  dplyr::select(1, 30:32, 22, 12, 25, 11, 6, 26, 13, 18, 4, 28, 17, 14, 8, 23, 9, 7, 16, 24, 2, 21, 5, 29, 10, 20, 19, 3, 15, 27)
```

Here, we show the resulting table for TRB:

```{r}
kable(excel_count_table_trb2 %>% head) %>%
  kable_styling(full_width = F, font_size = 11, 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```


# Prop tables and less than 5

Prop table TRA

```{r}

excel_count_table_tra3 <- excel_count_table_tra2 %>% mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no"))

excel_count_table_tra4 <- excel_count_table_tra3 %>% filter(nkt_trav11_traj18 == "no")

count_table_tra4 <- as.matrix(excel_count_table_tra4[,5:32])
rownames(count_table_tra4) <- excel_count_table_tra4$aaSeqCDR3

tra4_norm <- scale(count_table_tra4, center=FALSE, scale=colSums(count_table_tra4))

prop.table.tra <- cbind(tra4_norm, excel_count_table_tra4 %>% select(-starts_with("CD")) ) 

prop.table.tra2 <- prop.table.tra %>% dplyr::select(29, 31, 30, 32, 1:28)

prop.table.tra2 %>% head
ExpData(data=prop.table.tra2, type=2)

```

Prop table TRB

```{r}
excel_count_table_trb3 <- excel_count_table_trb2 %>% 
  mutate(nkt_trav11_traj18 = "no")

count_table_trb4 <- as.matrix(excel_count_table_trb3[,5:32])
rownames(count_table_trb4) <- excel_count_table_trb3$aaSeqCDR3


trb4_norm <- scale(count_table_trb4, center=FALSE, scale=colSums(count_table_trb4))

prop.table.trb <- cbind(trb4_norm, excel_count_table_trb %>% select(-starts_with("CD")) ) 
colnames(prop.table.trb)
prop.table.trb2 <- prop.table.trb %>% dplyr::select(29, 31, 30, 32, 1:28)


prop.table.trb2 %>% head
ExpData(data=prop.table.trb2, type=2)
```

# Gene segment usage analysis

```{r}
immdata_tra <- readRDS("immdata_tra.rds")
immdata_trb <- readRDS("immdata_trb.rds")
```



### TRA

```{r fig.width=25, fig.height=8}

### Thymus

cd4_thymus <- geneUsage(repFilter(immdata_tra, 
                                 .method = "by.meta", 
                                 .query = list(Organ = include('Thymus'), Cell_type = include('CD4')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trav")

vis(cd4_thymus, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_tra$meta) + ggtitle("TRA Thymus CD4") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))
#ggsave("./final_fig/gene_usage_new/cd4_thymus_tra_just1segment.png", width = 60, height = 12, units = "cm") 
#ggsave("./final_fig/gene_usage_new/cd4_thymus_tra_just1segment.svg", width = 60, height = 12, units = "cm") 


### LN
cd4_ln <- geneUsage(repFilter(immdata_tra, 
                                 .method = "by.meta", 
                                 .query = list(Organ = exclude('Thymus'), Cell_type = include('CD4')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trav")

vis(cd4_ln, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_tra$meta) + ggtitle("TRA LN CD4") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))


#ggsave("./final_fig/gene_usage_new/cd4_ln_tra_just1segment.png", width = 60, height = 12, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd4_ln_tra_just1segment.svg", width = 60, height = 12, units = "cm")

cd8_thymus <- geneUsage(repFilter(immdata_tra, 
                                 .method = "by.meta", 
                                 .query = list(Organ = include('Thymus'), Cell_type = include('CD8')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trav")

vis(cd8_thymus, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_tra$meta) + ggtitle("TRA Thymus CD8") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))
#ggsave("./final_fig/gene_usage_new/cd8_thymus_tra_just1segment.png", width = 60, height = 12, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd8_thymus_tra_just1segment.svg", width = 60, height = 12, units = "cm")


cd8_ln <- geneUsage(repFilter(immdata_tra, 
                                 .method = "by.meta", 
                                 .query = list(Organ = exclude('Thymus'), Cell_type = include('CD8')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trav")

vis(cd8_ln, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_tra$meta) + ggtitle("TRA LN CD8") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))

#ggsave("./final_fig/gene_usage_new/cd8_ln_tra_just1segment.png", width = 60, height = 12, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd8_ln_tra_just1segment.svg", width = 60, height = 12, units = "cm")


```

### TRB

```{r fig.width=8, fig.height=4}
### Thymus

cd4_thymus <- geneUsage(repFilter(immdata_trb, 
                                 .method = "by.meta", 
                                 .query = list(Organ = include('Thymus'), Cell_type = include('CD4')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trbv")

vis(cd4_thymus, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_trb$meta, .test = F) + ggtitle("TRB Thymus CD4") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70")) 
  
#ggsave("./final_fig/gene_usage_new/cd4_thymus_trb_just1segment.png", width = 20, height = 10, units = "cm") 
#ggsave("./final_fig/gene_usage_new/cd4_thymus_trb_just1segment.svg", width = 20, height = 10, units = "cm")

### LN
cd4_ln <- geneUsage(repFilter(immdata_trb, 
                                 .method = "by.meta", 
                                 .query = list(Organ = exclude('Thymus'), Cell_type = include('CD4')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trbv")

vis(cd4_ln, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_trb$meta, .test = F) + ggtitle("TRB LN CD4") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))

#ggsave("./final_fig/gene_usage_new/cd4_ln_trb_just1segment.png", width = 20, height = 10, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd4_ln_trb_just1segment.svg", width = 20, height = 10, units = "cm")


cd8_thymus <- geneUsage(repFilter(immdata_trb, 
                                 .method = "by.meta", 
                                 .query = list(Organ = include('Thymus'), Cell_type = include('CD8')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trbv")

vis(cd8_thymus, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_trb$meta, .test = F) + ggtitle("TRB Thymus CD8") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))
#ggsave("./final_fig/gene_usage_new/cd8_thymus_trb_just1segment.png", width = 20, height = 10, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd8_thymus_trb_just1segment.svg", width = 20, height = 10, units = "cm")


cd8_ln <- geneUsage(repFilter(immdata_trb, 
                                 .method = "by.meta", 
                                 .query = list(Organ = exclude('Thymus'), Cell_type = include('CD8')))$data, .norm = T, .ambig = "maj", .quant = "count", .gene = "musmus.trbv")

vis(cd8_ln, .by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_trb$meta, .test = F) + ggtitle("TRB LN CD8") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70"))

#ggsave("./final_fig/gene_usage_new/cd8_ln_trb_just1segment.png", width = 20, height = 10, units = "cm")
#ggsave("./final_fig/gene_usage_new/cd8_ln_trb_just1segment.svg", width = 20, height = 10, units = "cm")

```

# Analysis of NKT cells

Next, we will identify gene segments that are typical for invariant NKT cells. These segments are TRAV11 and TRAJ18 for TRA and TRBV1, TRBV13-2 and TRBV29 for TRB. Please, note that some of the included TRB chains are used by conventional cells as well. 

```{r}
# NKT analysis
merged_repertoires <- merged_repertoires %>% 
  mutate(is_nkt = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11")) & (grepl(allJHitsWithScore, pattern = "TRAJ18")) |
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")) |
      (grepl(allVHitsWithScore, pattern = "TRBV1\\*")) |
      (grepl(allVHitsWithScore, pattern = "TRBV29"))  ,"yes","no")) %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% mutate(nkt_trbv13_2 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")),"yes","no")) %>% mutate(nkt_trbv29 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV29")),"yes","no")) %>% mutate(nkt_trbv1 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV1\\*")),"yes","no"))

md <- md %>% select(Exp, Organ, Cell_type, Mouse_strain, num_id) %>% mutate_at(vars("num_id"), as.numeric)

# All TRB NKT
nkt_table_trb <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, is_nkt, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, is_nkt) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(is_nkt == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRAV11 TRAJ18
nkt_table_trav11_traj18 <- merged_repertoires %>% 
   filter(chain == "TRA") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trav11_traj18, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trav11_traj18) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trav11_traj18 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRBV1
nkt_table_trbv1 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv1, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv1) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv1 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRBV13-2
nkt_table_trbv13_2 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv13_2, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv13_2) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv13_2 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 


# TRBV29
nkt_table_trbv29 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv29, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv29) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv29 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table_trb$nkt_sample <- "all_nkt_trb"
nkt_table_trav11_traj18$nkt_sample <- "trav11_traj18"
nkt_table_trbv1$nkt_sample <- "trbv1"
nkt_table_trbv13_2$nkt_sample <- "trbv13_2"
nkt_table_trbv29$nkt_sample <- "trbv29"

nkt_table <- cbind(nkt_table_trb$num_id, nkt_table_trb$sample_type, nkt_table_trb$Exp, format(nkt_table_trb$freq, digits = 2), 
                   format(nkt_table_trav11_traj18$freq, digits = 2),
format(nkt_table_trbv1$freq, digits = 2), 
format(nkt_table_trbv13_2$freq, digits = 2),
                   format(nkt_table_trbv29$freq, digits = 2)) %>% as.data.frame

colnames(nkt_table) <- c("num_id","Sample","Exp","All TRB NKT", "TRAV11-TRAJ18", "TRBV1","TRBV13-2","TRBV29")
nkt_table <- nkt_table %>% mutate(Sample = paste(Sample, Exp)) %>% select(-num_id, -Exp) %>% arrange(Sample)
```

See the percentage of NKT gene segments in each sample:

```{r}
kable(nkt_table, format = "html") %>%
  kable_styling(full_width = F, font_size = 11, 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


Now we will plot the data for CD4 cells and for CD8 cells. Plots for TRA and TRB sequences are separated. 

## All NKT

This plot visualizes the percentage of any of the TRB NKT chains (TRBV1, TRBV13-2 and TRBV29). We can see that this distinction is not specific as these gene segments are also used by conventional CD4 and CD8 T cells. 

```{r fig.width=7, fig.height=3.5}
## CD4
p01 <- nkt_table_trb %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("NKT segments in CD4 TRB") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p02 <- nkt_table_trb %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("NKT segments in CD8 TRB") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.eps", width = 8, height = 8, units = "cm")

p01 + p02
```


### TRAV11 TRAJ18

```{r fig.width=7, fig.height=3.5}
## CD4
p03 <- nkt_table_trav11_traj18 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRAV11-TRAJ18 segments in CD4 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p04 <- nkt_table_trav11_traj18 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRAV11-TRAJ18 segments in CD8 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,1)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.eps", width = 8, height = 8, units = "cm")

p03 + p04
```

### TRBV1

```{r fig.width=7, fig.height=3.5}

## CD4
p05 <- nkt_table_trbv1 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV1") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p05 <- nkt_table_trbv1 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV1") + 
  theme_classic() + 
  ggtheme() + 
  
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.eps", width = 8, height = 8, units = "cm")

p05 + p06
```

### TRBV13-2

```{r fig.width=7, fig.height=3.5}

## CD4
p07 <- nkt_table_trbv13_2 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRBV13-2 segments in CD4 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p08 <- nkt_table_trbv13_2 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRBV13-2 segments in CD8 cells") + 
  theme_classic() + 
  ggtheme() + 
  
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.eps", width = 8, height = 8, units = "cm")

p07 + p08
```

### TRBV29

```{r fig.width=7, fig.height=3.5}

## CD4
p09 <- nkt_table_trbv29 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV29") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p10 <- nkt_table_trbv29 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV29") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.eps", width = 8, height = 8, units = "cm")

p09 + p10
```

# PCA 

In this part we explore the similarity of TCR repertoires of different samples using principal component analysis (PCA).

We will calculate the distances from a filtered matrix of proportions, in which we filtered out the CDR3 sequences that are present in less than five CD4 or in less than five CD8 samples. 

First, we will prepare the filtered normalized count matrix. 

## less than 5 removed

## CD4

Remove CDR3s with occurrence in less than 5 samples

```{r}

# TRA
keep_tra_cd4 <- excel_count_table_tra %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") | 
    (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% 
  filter(nkt_trav11_traj18 == "no") %>% 
  select(aaSeqCDR3, starts_with("CD4")) %>% 
  mutate_at(vars(starts_with("CD4")), .funs = binary) 
 
keep_tra_cd4$sum <- rowSums((keep_tra_cd4 %>% select(-aaSeqCDR3)))

keep_tra_cd4$keep <- ifelse(keep_tra_cd4$sum>4,1,0)

keep_tra_cd4_sequences <- pull(keep_tra_cd4 %>% filter(keep == 1), aaSeqCDR3)

# TRB
keep_trb_cd4 <- excel_count_table_trb %>% select(aaSeqCDR3, starts_with("CD4")) %>%  mutate_at(vars(starts_with("CD4")), .funs = binary) 
 
keep_trb_cd4$sum <- rowSums((keep_trb_cd4 %>% select(-aaSeqCDR3)))

keep_trb_cd4$keep <- ifelse(keep_trb_cd4$sum>4,1,0)

keep_trb_cd4_sequences <- pull(keep_trb_cd4 %>% filter(keep == 1), aaSeqCDR3)

```

## CD8

Remove CDR3s with occurrence in less than 5 samples

```{r}

# TRA
keep_tra_cd8 <- excel_count_table_tra %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") | 
    (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% 
  filter(nkt_trav11_traj18 == "no") %>% 
  select(aaSeqCDR3, starts_with("CD8")) %>% 
  mutate_at(vars(starts_with("CD8")), .funs = binary) 

keep_tra_cd8$sum <- rowSums((keep_tra_cd8 %>% select(-aaSeqCDR3)))

keep_tra_cd8$keep <- ifelse(keep_tra_cd8$sum>4,1,0)

keep_tra_cd8_sequences <- pull(keep_tra_cd8 %>% filter(keep == 1), aaSeqCDR3)

# TRB
keep_trb_cd8 <- excel_count_table_trb %>% select(aaSeqCDR3, starts_with("CD8")) %>%  mutate_at(vars(starts_with("CD8")), .funs = binary) 
 
keep_trb_cd8$sum <- rowSums((keep_trb_cd8 %>% select(-aaSeqCDR3)))

keep_trb_cd8$keep <- ifelse(keep_trb_cd8$sum>4,1,0)

keep_trb_cd8_sequences <- pull(keep_trb_cd8 %>% filter(keep == 1), aaSeqCDR3)

```

## TRB

```{r}
excel_count_table_trb3_filter <- excel_count_table_trb2  %>% filter(aaSeqCDR3 %in% keep_trb_cd4_sequences | aaSeqCDR3 %in% keep_trb_cd8_sequences) 

count_table_trb4_filter <- as.matrix(excel_count_table_trb3_filter[,5:32])
rownames(count_table_trb4_filter) <- excel_count_table_trb3_filter$aaSeqCDR3

trb4_norm_filter <- scale(count_table_trb4_filter, center=FALSE, scale=colSums(count_table_trb4_filter))

prop.table.trb_filter <- cbind(trb4_norm_filter, excel_count_table_trb3_filter %>% select(-starts_with("CD")) )
```


## TRA

```{r}
excel_count_table_tra3_filter <- excel_count_table_tra3  %>% filter(aaSeqCDR3 %in% keep_tra_cd4_sequences | aaSeqCDR3 %in% keep_tra_cd8_sequences) %>% filter(nkt_trav11_traj18 == "no")

count_table_tra4_filter <- as.matrix(excel_count_table_tra3_filter[,5:32])
rownames(count_table_tra4_filter) <- excel_count_table_tra3_filter$aaSeqCDR3

tra4_norm_filter <- scale(count_table_tra4_filter, center=FALSE, scale=colSums(count_table_tra4_filter))

prop.table.tra_filter <- cbind(tra4_norm_filter, excel_count_table_tra3_filter %>% select(-starts_with("CD")) )

colnames(prop.table.tra_filter)
```


```{r}
prop.table.tra2_filter.cd4 <- prop.table.tra_filter[,1:14]
prop.table.tra2_filter.cd4$sum <- rowSums(prop.table.tra2_filter.cd4)
prop.table.tra2_filter.cd4 <- prop.table.tra2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

prop.table.trb2_filter.cd4 <- prop.table.trb_filter[,1:14]
prop.table.trb2_filter.cd4$sum <- rowSums(prop.table.trb2_filter.cd4)
prop.table.trb2_filter.cd4 <- prop.table.trb2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

prop.table.tra2_filter.cd8 <- prop.table.tra_filter[,15:28]
prop.table.tra2_filter.cd8$sum <- rowSums(prop.table.tra2_filter.cd8)
prop.table.tra2_filter.cd8 <- prop.table.tra2_filter.cd8 %>% filter(sum >0) %>% select(-sum)

prop.table.trb2_filter.cd8 <- prop.table.trb_filter[,15:28]
prop.table.trb2_filter.cd8$sum <- rowSums(prop.table.trb2_filter.cd8)
prop.table.trb2_filter.cd8 <- prop.table.trb2_filter.cd8 %>% filter(sum >0) %>% select(-sum)
```


## TRA + TRB together

### CD4

```{r fig.width=4, fig.height=3}
prop.table.tra2_filter.merge <- rbind(prop.table.tra2_filter.cd4, prop.table.trb2_filter.cd4)
  
res.pca.merge.cd4 <- prcomp(t(prop.table.tra2_filter.merge), scale = TRUE, center = T)

mdres.pca.merge.cd4  <-  colnames(prop.table.tra2_filter.merge[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd4$Organ),
                  color = as.factor(mdres.pca.merge.cd4$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  ggtheme() +
  theme(axis.text.x = element_text(angle = 0))

#ggsave("final_fig/pca/cd4.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4.svg", width = 2.6, height = 1.7)
```

Zoom 

```{r fig.width=4, fig.height=3}
fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd4$Organ),
                  color = as.factor(mdres.pca.merge.cd4$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  xlim(c(-30,0)) + ylim(-2,15) +
  ggtheme() +
  theme(axis.text.x = element_text(angle = 0))


#ggsave("final_fig/pca/cd4_zoom.png", width = 2.2, height = 1.4)
#ggsave("final_fig/pca/cd4_zoom.svg", width = 2.2, height = 1.4)
```


### CD8

```{r fig.width=4, fig.height=3}
prop.table.tra2_filter.merge <- rbind(prop.table.tra2_filter.cd8, prop.table.trb2_filter.cd8)
  
res.pca.merge.cd8 <- prcomp(t(prop.table.tra2_filter.merge), scale = TRUE, center = T)

mdres.pca.merge.cd8  <-  colnames(prop.table.tra2_filter.merge[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.merge.cd8,
             col.ind = as.factor(mdres.pca.merge.cd8$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd8$Organ),
                  color = as.factor(mdres.pca.merge.cd8$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  ggtheme()
#ggsave("final_fig/pca/cd8.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8.svg", width = 2.6, height = 1.7)
```

Zoom

```{r fig.width=4, fig.height=3}
fviz_pca_ind(res.pca.merge.cd8,
             col.ind = as.factor(mdres.pca.merge.cd8$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd8$Organ),
                  color = as.factor(mdres.pca.merge.cd8$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  xlim(c(-32,-18)) + ylim(-4,5) +
  ggtheme()

#ggsave("final_fig/pca/cd8_zoom.png", width = 2.2, height = 1.4)
#ggsave("final_fig/pca/cd8_zoom.svg", width = 2.2, height = 1.4)
```

## TRA

```{r fig.width=4, fig.height=3}
res.pca <- prcomp(t(prop.table.tra_filter[,1:28]), scale = TRUE)


md  <-  colnames(prop.table.tra_filter[,1:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

```

### CD4

```{r fig.width=4, fig.height=3}
## All

prop.table.tra2_filter.cd4 <- prop.table.tra_filter[,1:14]
prop.table.tra2_filter.cd4$sum <- rowSums(prop.table.tra2_filter.cd4)

prop.table.tra2_filter.cd4 <- prop.table.tra2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4 <- prcomp(t(prop.table.tra2_filter.cd4), scale = TRUE, center = T)

mdres.pca.tra.cd4  <-  colnames(prop.table.tra_filter[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
# LN
prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd4.ln <- prop.table.tra_filter[,1:7]
prop.table.tra2_filter.cd4.ln$sum <- rowSums(prop.table.tra2_filter.cd4.ln)

prop.table.tra2_filter.cd4.ln <- prop.table.tra2_filter.cd4.ln %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4.ln <- prcomp(t(prop.table.tra2_filter.cd4.ln), scale = TRUE, center = T)

mdres.pca.tra.cd4.ln  <-  colnames(prop.table.tra_filter[,1:7])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd4.ln,
             col.ind = as.factor(mdres.pca.tra.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd4.ln$Organ),
                  color = as.factor(mdres.pca.tra.cd4.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_tra_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_tra_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.tra2_filter.cd4.thy <- prop.table.tra_filter[,8:14]
prop.table.tra2_filter.cd4.thy$sum <- rowSums(prop.table.tra2_filter.cd4.thy)

prop.table.tra2_filter.cd4.thy <- prop.table.tra2_filter.cd4.thy %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4.thy <- prcomp(t(prop.table.tra2_filter.cd4.thy), scale = TRUE, center = T)

mdres.pca.tra.cd4.thy  <-  colnames(prop.table.tra_filter[,8:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
fviz_pca_ind(res.pca.tra.cd4.thy,
             col.ind = as.factor(mdres.pca.tra.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd4.thy$Organ),
                  color = as.factor(mdres.pca.tra.cd4.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") +
  ggtheme()
#ggsave("final_fig/pca/cd4_tra_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_tra_thy.svg", width = 2.6, height = 1.7)

```

### CD8

```{r fig.width=4, fig.height=3}
## All

prop.table.tra2_filter.cd8 <- prop.table.tra_filter[,15:28]
prop.table.tra2_filter.cd8$sum <- rowSums(prop.table.tra2_filter.cd8)

prop.table.tra2_filter.cd8 <- prop.table.tra2_filter.cd8 %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8 <- prcomp(t(prop.table.tra2_filter.cd8), scale = TRUE, center = T)

mdres.pca.tra.cd8  <-  colnames(prop.table.tra2_filter[,15:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

# LN
prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd8.ln <- prop.table.tra_filter[,15:21]
prop.table.tra2_filter.cd8.ln$sum <- rowSums(prop.table.tra2_filter.cd8.ln)

prop.table.tra2_filter.cd8.ln <- prop.table.tra2_filter.cd8.ln %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8.ln <- prcomp(t(prop.table.tra2_filter.cd8.ln), scale = TRUE, center = T)

mdres.pca.tra.cd8.ln  <-  colnames(prop.table.tra_filter[,15:21])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
fviz_pca_ind(res.pca.tra.cd8.ln,
             col.ind = as.factor(mdres.pca.tra.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd8.ln$Organ),
                  color = as.factor(mdres.pca.tra.cd8.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_tra_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_tra_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd8.thy <- prop.table.tra_filter[,22:28]
prop.table.tra2_filter.cd8.thy$sum <- rowSums(prop.table.tra2_filter.cd8.thy)

prop.table.tra2_filter.cd8.thy <- prop.table.tra2_filter.cd8.thy %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8.thy <- prcomp(t(prop.table.tra2_filter.cd8.thy), scale = TRUE, center = T)

mdres.pca.tra.cd8.thy  <-  colnames(prop.table.tra_filter[,22:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
 
fviz_pca_ind(res.pca.tra.cd8.thy,
             col.ind = as.factor(mdres.pca.tra.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd8.thy$Organ),
                  color = as.factor(mdres.pca.tra.cd8.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_tra_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_tra_thy.svg", width = 2.6, height = 1.7)

```

## TRB

```{r fig.width=4, fig.height=3}
res.pca.trb <- prcomp(t(prop.table.trb_filter[,5:32]), scale = TRUE)


md_trb  <-  colnames(prop.table.trb_filter[,5:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))

```

### CD4

```{r fig.width=4, fig.height=3}
## All
colnames(prop.table.trb2_filter)

prop.table.trb2_filter.cd4 <- prop.table.trb_filter[,5:18]
prop.table.trb2_filter.cd4$sum <- rowSums(prop.table.trb2_filter.cd4)

prop.table.trb2_filter.cd4 <- prop.table.trb2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4 <- prcomp(t(prop.table.trb2_filter.cd4), scale = TRUE, center = T)

mdres.pca.trb.cd4  <-  colnames(prop.table.trb_filter[,5:18])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
 

# LN
prop.table.trb_filter %>% colnames
prop.table.trb2_filter.cd4.ln <- prop.table.trb_filter[,5:11]
prop.table.trb2_filter.cd4.ln$sum <- rowSums(prop.table.trb2_filter.cd4.ln)

prop.table.trb2_filter.cd4.ln <- prop.table.trb2_filter.cd4.ln %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4.ln <- prcomp(t(prop.table.trb2_filter.cd4.ln), scale = TRUE, center = T)

mdres.pca.trb.cd4.ln  <-  colnames(prop.table.trb_filter[,5:11])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd4.ln,
             col.ind = as.factor(mdres.pca.trb.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd4.ln$Organ),
                  color = as.factor(mdres.pca.trb.cd4.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_trb_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_trb_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.trb_filter %>% colnames
prop.table.trb2_filter.cd4.thy <- prop.table.trb_filter[,12:18]
prop.table.trb2_filter.cd4.thy$sum <- rowSums(prop.table.trb2_filter.cd4.thy)

prop.table.trb2_filter.cd4.thy <- prop.table.trb2_filter.cd4.thy %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4.thy <- prcomp(t(prop.table.trb2_filter.cd4.thy), scale = TRUE, center = T)

mdres.pca.trb.cd4.thy  <-  colnames(prop.table.trb_filter[,12:18])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
 
fviz_pca_ind(res.pca.trb.cd4.thy,
             col.ind = as.factor(mdres.pca.trb.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd4.thy$Organ),
                  color = as.factor(mdres.pca.trb.cd4.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_trb_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_trb_thy.svg", width = 2.6, height = 1.7)

```

### CD8

```{r fig.width=4, fig.height=3}
## All

prop.table.trb2_filter.cd8 <- prop.table.trb_filter[,19:32]
prop.table.trb2_filter.cd8$sum <- rowSums(prop.table.trb2_filter.cd8)

prop.table.trb2_filter.cd8 <- prop.table.trb2_filter.cd8 %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8 <- prcomp(t(prop.table.trb2_filter.cd8), scale = TRUE, center = T)

mdres.pca.trb.cd8  <-  colnames(prop.table.trb_filter[,19:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

# LN
prop.table.trb_filter %>% colnames
prop.table.trb2_filter.cd8.ln <- prop.table.trb_filter[,19:25]
prop.table.trb2_filter.cd8.ln$sum <- rowSums(prop.table.trb2_filter.cd8.ln)

prop.table.trb2_filter.cd8.ln <- prop.table.trb2_filter.cd8.ln %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8.ln <- prcomp(t(prop.table.trb2_filter.cd8.ln), scale = TRUE, center = T)

mdres.pca.trb.cd8.ln  <-  colnames(prop.table.trb_filter[,19:25])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
fviz_pca_ind(res.pca.trb.cd8.ln,
             col.ind = as.factor(mdres.pca.trb.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd8.ln$Organ),
                  color = as.factor(mdres.pca.trb.cd8.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_trb_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_trb_ln.svg", width = 2.6, height = 1.7)


# Thymus

prop.table.trb_filter %>% colnames
prop.table.trb2_filter.cd8.thy <- prop.table.trb_filter[,26:32]
prop.table.trb2_filter.cd8.thy$sum <- rowSums(prop.table.trb2_filter.cd8.thy)

prop.table.trb2_filter.cd8.thy <- prop.table.trb2_filter.cd8.thy %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8.thy <- prcomp(t(prop.table.trb2_filter.cd8.thy), scale = TRUE, center = T)

mdres.pca.trb.cd8.thy  <-  colnames(prop.table.trb_filter[,26:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
 
fviz_pca_ind(res.pca.trb.cd8.thy,
             col.ind = as.factor(mdres.pca.trb.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd8.thy$Organ),
                  color = as.factor(mdres.pca.trb.cd8.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_trb_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_trb_thy.svg", width = 2.6, height = 1.7)

```

# Diversity index

```{r}

immdata_tra_wonkt <- repFilter(immdata_tra, "by.clonotype",
  list(V.name = exclude("TRAV11"), J.name = exclude("TRAJ18")),
  .match = "substring")

## ALL TRA
repDiversity(immdata_tra_wonkt$data) %>% vis(.by = c( "Mouse_strain"), .meta = immdata_tra$meta, 
  .signif.label.size = 0) + ggtitle("TRA without NKT") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70")) + ggtitle("all TRA", subtitle = "")
#ggsave("./final_fig/diversity/chao_tra_wonkt_sign.png", width = 8, height = 13, units = "cm")
#ggsave("./final_fig/diversity/chao_tra_wonkt_sign.svg", width = 8, height = 13, units = "cm")


## ALL TRB
repDiversity(immdata_trb$data) %>% vis(.by = c( "Mouse_strain"), .meta = immdata_trb$meta, 
  .signif.label.size = 0) + ggtitle("TRB without NKT") + scale_fill_manual(values = c("dodgerblue","indianred2","gray70")) + ggtitle("all TRB", subtitle = "")
#ggsave("./final_fig/diversity/chao_trb_sign.png", width = 8, height = 13, units = "cm")
#ggsave("./final_fig/diversity/chao_trb_sign.svg", width = 8, height = 13, units = "cm")


## Chao index
repDiversity(immdata_tra_wonkt$data) %>% vis(.by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_tra$meta, .test = F,
  .signif.label.size = 0) + ggtitle("TRA")
#ggsave("./plots/chao_tra.png", width = 15, height = 11, units = "cm")
#ggsave("./plots/chao_tra.svg", width = 15, height = 11, units = "cm")

repDiversity(immdata_trb$data) %>% vis(.by = c("Cell_type", "Organ","Mouse_strain"), .meta = immdata_trb$meta, .test = F,
  .signif.label.size = 0) + ggtitle("TRB", subtitle = "")
#ggsave("./plots/chao_trb.png", width = 15, height = 11, units = "cm")
#ggsave("./plots/chao_trb.svg", width = 15, height = 11, units = "cm")

```

```{r}
chao_tra <- as.data.frame(repDiversity(immdata_tra_wonkt$data))
chao_trb <- as.data.frame(repDiversity(immdata_trb$data))
```


```{r}

chao_tra2 <- chao_tra %>% 
  rownames_to_column("num_id") %>% 
mutate(num_id = str_replace_all(string = num_id, c("new_lib_24" = "lib24_S24_L001",  "new_lib_25" = "lib25_S25_L001"))) %>% 
  separate(num_id, into = c("num_id",NA,NA), sep = "_") %>% 
  mutate(num_id = str_replace(num_id, "lib","")) %>% 
  mutate_at(vars("num_id"), as.numeric) 
  

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

# Attach metadata sent for sequencing

chao_tra2 <- left_join(chao_tra2, md) %>% select(value = Estimator, Cell_type, Organ, Mouse_strain) %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

chao_tra2 %>% filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = value, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Chao index estimate") + 
  ggtitle("CD4 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_cd4_tra.png", width = 8, height = 10, units = "cm")
#ggsave("./final_fig/diversity/chao_cd4_tra.eps", width = 8, height = 10, units = "cm")

levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

chao_tra2 %>% filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = value, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Chao index estimate") + 
  ggtitle("CD8 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_cd8_tra.png", width = 8, height = 10, units = "cm")
#ggsave("./final_fig/diversity/chao_cd8_tra.eps", width = 8, height = 10, units = "cm")

```

```{r}
chao_trb2 <- chao_trb %>% 
  rownames_to_column("num_id") %>% mutate(num_id = str_replace_all(string = num_id, c("new_lib_24" = "lib24_S24_L001",  "new_lib_25" = "lib25_S25_L001"))) %>% 
  separate(num_id, into = c("num_id",NA,NA), sep = "_") %>% 
  mutate(num_id = str_replace(num_id, "lib",""))  %>% 
  mutate_at(vars("num_id"), as.numeric) 
  

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

chao_trb2 <- left_join(chao_trb2, md) %>% select(value = Estimator, Cell_type, Organ, Mouse_strain) %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

chao_trb2 %>% filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = value, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Chao index estimate") + 
  ggtitle("CD4 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,125500)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_cd4_trb.png", width = 8, height = 10, units = "cm")
#ggsave("./final_fig/diversity/chao_cd4_trb.eps", width = 8, height = 10, units = "cm")

levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")


chao_trb2 %>% filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = value, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Chao index estimate") + 
  ggtitle("CD8 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,125500)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_cd8_trb.png", width = 8, height = 10, units = "cm")
#ggsave("./final_fig/diversity/chao_cd8_trb.eps", width = 8, height = 10, units = "cm")

```


```{r}
chao_trb2 %>% 
  ggplot(aes(y = value, x = factor(Mouse_strain, levels = c("WT","CA","CAKR")))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.1), size = 2, aes(color = Mouse_strain)) +
ylab("Chao index estimate") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_all_trb.png", width = 8, height = 6, units = "cm")
#ggsave("./final_fig/diversity/chao_all_trb.eps", width = 8, height = 6, units = "cm")


chao_tra2 %>% 
  ggplot(aes(y = value, x = factor(Mouse_strain, levels = c("WT","CA","CAKR")))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.1), size = 2, aes(color = Mouse_strain)) +
ylab("Chao index estimate") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/diversity/chao_all_tra.png", width = 8, height = 6, units = "cm")
#ggsave("./final_fig/diversity/chao_all_tra.eps", width = 8, height = 6, units = "cm")

```

# Top20 WT LN sequences

```{r}
topclones <- prop.table.trb_filter %>% mutate(topclones_ln_cd4 = (`CD4 Lymph nodes WT Exp02`+
                                                                       `CD4 Lymph nodes WT Exp03`)/2,
                                               topclones_ln_cd8 = (`CD8 Lymph nodes WT Exp01`+
                                                                       `CD8 Lymph nodes WT Exp02`+
                                                                       `CD8 Lymph nodes WT Exp03`)/3)

cd4_topclones_20sequences_trb <- pull(topclones %>% slice_max(order_by = topclones_ln_cd4, n = 20), aaSeqCDR3)
cd8_topclones_20sequences_trb <- pull(topclones %>% slice_max(order_by = topclones_ln_cd8, n = 20), aaSeqCDR3)
```

```{r}
topclones <- prop.table.tra_filter %>% mutate(topclones_ln_cd4 = (`CD4 Lymph nodes WT Exp02`+
                                                                       `CD4 Lymph nodes WT Exp03`)/2,
                                               topclones_ln_cd8 = (`CD8 Lymph nodes WT Exp01`+
                                                                       `CD8 Lymph nodes WT Exp02`+
                                                                       `CD8 Lymph nodes WT Exp03`)/3)

cd4_topclones_20sequences_tra <- pull(topclones %>% slice_max(order_by = topclones_ln_cd4, n = 20), aaSeqCDR3)
cd8_topclones_20sequences_tra <- pull(topclones %>% slice_max(order_by = topclones_ln_cd8, n = 20), aaSeqCDR3)

```


## Summary plot

### CD4 TRA

```{r fig.width=3, fig.height=3}
cd4_topclones_rep_pct <- prop.table.tra2 %>% filter(aaSeqCDR3 %in% cd4_topclones_20sequences_tra) %>% select(aaSeqCDR3, starts_with("CD4")) %>% select(1,14,15,9:13,7,8,2:6) 

cd4_topclones_rep_pct2 <- as.data.frame(colSums(cd4_topclones_rep_pct[,2:15]))
colnames(cd4_topclones_rep_pct2) <- "value"

cd4_topclones_rep_pct2 <- cd4_topclones_rep_pct2 %>% 
  rownames_to_column("sample") %>% 
  mutate(sample = stringr::str_replace_all(sample, pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp")) %>% 
  mutate(Cell_type_organ = factor(paste(Organ, Strain), levels = c("Thymus WT","Thymus CA", "Thymus CAKR",
                                                                   "LN WT","LN CA", "LN CAKR" )))

cd4_topclones_rep_pct2

cd4_topclones_rep_pct2 %>% 
  filter(Organ == "LN") %>% 
  ggplot(aes(y = value*100, x = factor(Cell_type_organ))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Strain)) + 
  ylab("% of repertoire") + 
  ggtitle("CD4 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")

#ggsave("./final_fig/20topseqs/20topseqs_pct_tra_cd4.png", width = 5, height = 5.7, units = "cm")
#ggsave("./final_fig/20topseqs/20topseqs_pct_tra_cd4.eps", width = 5, height = 5.7, units = "cm")

```

### CD4 TRB
```{r fig.width=3, fig.height=3}
cd4_topclones_rep_pct <- prop.table.trb2 %>% filter(aaSeqCDR3 %in% cd4_topclones_20sequences_trb) %>% select(aaSeqCDR3, starts_with("CD4")) %>% select(1,14,15,9:13,7,8,2:6) 

cd4_topclones_rep_pct2 <- as.data.frame(colSums(cd4_topclones_rep_pct[,2:15]))
colnames(cd4_topclones_rep_pct2) <- "value"

cd4_topclones_rep_pct2 <- cd4_topclones_rep_pct2 %>% 
  rownames_to_column("sample") %>% 
  mutate(sample = stringr::str_replace_all(sample, pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp")) %>% 
  mutate(Cell_type_organ = factor(paste(Organ, Strain), levels = c("Thymus WT","Thymus CA", "Thymus CAKR",
                                                                   "LN WT","LN CA", "LN CAKR" )))

cd4_topclones_rep_pct2

cd4_topclones_rep_pct2 %>% 
  filter(Organ == "LN") %>% 
  ggplot(aes(y = value*100, x = factor(Cell_type_organ))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Strain)) + 
  ylab("% of repertoire") + 
  ggtitle("CD4 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")


#ggsave("./final_fig/20topseqs/20topseqs_pct_trb_cd4.png", width = 5.4, height = 5.7, units = "cm")
#ggsave("./final_fig/20topseqs/20topseqs_pct_trb_cd4.eps", width = 5.4, height = 5.7, units = "cm")



```

### CD8 TRA

```{r fig.width=3, fig.height=3}
cd8_topclones_rep_pct <- prop.table.tra2 %>% filter(aaSeqCDR3 %in% cd8_topclones_20sequences_tra) %>% select(aaSeqCDR3, starts_with("CD8")) %>% select(1,13:15,9:12,6:8,2:5) 

cd8_topclones_rep_pct2 <- as.data.frame(colSums(cd8_topclones_rep_pct[,2:15]))
colnames(cd8_topclones_rep_pct2) <- "value"

cd8_topclones_rep_pct2 <- cd8_topclones_rep_pct2 %>% 
  rownames_to_column("sample") %>% 
  mutate(sample = stringr::str_replace_all(sample, pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp")) %>% 
  mutate(Cell_type_organ = factor(paste(Organ, Strain), levels = c("Thymus WT","Thymus CA", "Thymus CAKR",
                                                                   "LN WT","LN CA", "LN CAKR" )))

cd8_topclones_rep_pct2

cd8_topclones_rep_pct2 %>% 
  filter(Organ == "LN") %>% 
  ggplot(aes(y = value*100, x = factor(Cell_type_organ))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Strain)) + 
  ylab("% of repertoire") + 
  ggtitle("CD8 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")

#ggsave("./final_fig/20topseqs/20topseqs_pct_tra_cd8.png", width = 5, height = 5.7, units = "cm")
#ggsave("./final_fig/20topseqs/20topseqs_pct_tra_cd8.eps", width = 5, height = 5.7, units = "cm")

```

### CD8 TRB
```{r fig.width=3, fig.height=3}
cd8_topclones_rep_pct <- prop.table.trb2 %>% filter(aaSeqCDR3 %in% cd8_topclones_20sequences_trb) %>% select(aaSeqCDR3, starts_with("CD8")) %>% select(1,13:15,9:12,6:8,2:5) 

cd8_topclones_rep_pct2 <- as.data.frame(colSums(cd8_topclones_rep_pct[,2:15]))
colnames(cd8_topclones_rep_pct2) <- "value"

cd8_topclones_rep_pct2 <- cd8_topclones_rep_pct2 %>% 
  rownames_to_column("sample") %>% 
  mutate(sample = stringr::str_replace_all(sample, pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp")) %>% 
  mutate(Cell_type_organ = factor(paste(Organ, Strain), levels = c("Thymus WT","Thymus CA", "Thymus CAKR",
                                                                   "LN WT","LN CA", "LN CAKR" )))

cd8_topclones_rep_pct2

cd8_topclones_rep_pct2 %>% 
  filter(Organ == "LN") %>% 
  ggplot(aes(y = value*100, x = factor(Cell_type_organ))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Strain)) + 
  ylab("% of repertoire") + 
  ggtitle("CD8 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")

#ggsave("./final_fig/20topseqs/20topseqs_pct_trb_cd8.png", width = 5.4, height = 5.7, units = "cm")
#ggsave("./final_fig/20topseqs/20topseqs_pct_trb_cd8.eps", width = 5.4, height = 5.7, units = "cm")

```

# Heatmaps

## TRA

### CD4

```{r fig.width=6, fig.height=8}

# TRA
cd4_topclones_heatmap_matrix <- prop.table.tra2 %>% filter(aaSeqCDR3 %in% cd4_topclones_20sequences_tra) %>% select(aaSeqCDR3, starts_with("CD4")) %>% select(1,14,15,9:13,7,8,2:6)

cd4_topclones_heatmap_matrix2 <- as.matrix(cd4_topclones_heatmap_matrix[,2:15])
rownames(cd4_topclones_heatmap_matrix2) <- cd4_topclones_heatmap_matrix$aaSeqCDR3

cd4_topclones_heatmap_matrix2 <- cd4_topclones_heatmap_matrix2[match( cd4_topclones_20sequences_tra, rownames(cd4_topclones_heatmap_matrix2)),]

pheatmap(cd4_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)

#ggsave(plot = pheatmap(cd4_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F), "final_fig/heatmap_top20_ordered/cd4_tra_noscale.svg")


```

### CD8 

```{r fig.width=6, fig.height=8}
# TRA

cd8_topclones_heatmap_matrix <- prop.table.tra2 %>% filter(aaSeqCDR3 %in% cd8_topclones_20sequences_tra) %>% select(aaSeqCDR3, starts_with("CD8")) %>% select(1,13:15,9:12,6:8,2:5)
  

cd8_topclones_heatmap_matrix2 <- as.matrix(cd8_topclones_heatmap_matrix[,2:15])
rownames(cd8_topclones_heatmap_matrix2) <- cd8_topclones_heatmap_matrix$aaSeqCDR3

cd8_topclones_heatmap_matrix2 <- cd8_topclones_heatmap_matrix2[match( cd8_topclones_20sequences_tra, rownames(cd8_topclones_heatmap_matrix2)),]

pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)

#ggsave(plot = pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F), "final_fig/heatmap_top20_ordered/cd8_tra_noscale.svg")
```

without 1st sequence
```{r fig.width=6, fig.height=8}
### without 1st sequence

cd8_topclones_heatmap_matrix2 <- as.matrix(cd8_topclones_heatmap_matrix[2:20,2:15])
rownames(cd8_topclones_heatmap_matrix2) <- cd8_topclones_heatmap_matrix$aaSeqCDR3[2:20]

cd8_topclones_heatmap_matrix2 <- cd8_topclones_heatmap_matrix2[match( cd8_topclones_20sequences_tra[2:20], rownames(cd8_topclones_heatmap_matrix2)),]

pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)

#ggsave(plot = pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F), "final_fig/heatmap_top20_ordered/cd8_tra_noscale_without.svg")
```

## TRB

### CD4

```{r fig.width=6, fig.height=8}

# trb

cd4_topclones_heatmap_matrix <- prop.table.trb2 %>% filter(aaSeqCDR3 %in% cd4_topclones_20sequences_trb) %>% select(aaSeqCDR3, starts_with("CD4")) %>% select(1,14,15,9:13,7,8,2:6)
  
cd4_topclones_heatmap_matrix2 <- as.matrix(cd4_topclones_heatmap_matrix[,2:15])
rownames(cd4_topclones_heatmap_matrix2) <- cd4_topclones_heatmap_matrix$aaSeqCDR3

cd4_topclones_heatmap_matrix2 <- cd4_topclones_heatmap_matrix2[match( cd4_topclones_20sequences_trb, rownames(cd4_topclones_heatmap_matrix2)),]

pheatmap(cd4_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)

#ggsave(plot = pheatmap(cd4_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F), "final_fig/heatmap_top20_ordered/cd4_trb_noscale.svg")
```

### CD8

```{r fig.width=6, fig.height=8}
# trb

cd8_topclones_heatmap_matrix <- prop.table.trb2 %>% filter(aaSeqCDR3 %in% cd8_topclones_20sequences_trb) %>% select(aaSeqCDR3, starts_with("CD8")) %>% select(1,13:15,9:12,6:8,2:5)
  
cd8_topclones_heatmap_matrix2 <- as.matrix(cd8_topclones_heatmap_matrix[,2:15])
rownames(cd8_topclones_heatmap_matrix2) <- cd8_topclones_heatmap_matrix$aaSeqCDR3

cd8_topclones_heatmap_matrix2 <- cd8_topclones_heatmap_matrix2[match( cd8_topclones_20sequences_trb, rownames(cd8_topclones_heatmap_matrix2)),]

pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)

# png("final_fig/heatmap_top20_ordered/cd8_trb_noscale.png", width = 400, height = 500)
# pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F)
# dev.off()

#ggsave(plot = pheatmap(cd8_topclones_heatmap_matrix2, scale = "none", cluster_rows = F, cluster_cols = F), "final_fig/heatmap_top20_ordered/cd8_trb_noscale.svg")
```

# Repertoire overlap

```{r}

binary_tra <- excel_count_table_tra %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") | 
    (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% 
  filter(nkt_trav11_traj18 == "no") %>% 
  mutate_at(vars(starts_with("CD")), .funs = binary) 

# count_overlap <- function(df, column_number){
#   repertoire <- df %>% select(vars(column_number), aaSeqCDR3) %>% filter(vars(column_number)>0) %>% pull("aaSeqCDR3")
#   nrow_column <- length(repertoire)
#   intersect_rept <- length(intersect(repertoire, most_diverse_repertoire))
#   pct <- intersect_rept/length()
#   return(x)
# }

binary_tra_longer <- binary_tra %>% select(starts_with("CD"), aaSeqCDR3) %>% pivot_longer(!aaSeqCDR3, names_to = "num_id")

df_all4 <- data.frame("")

for(j in levels(factor(binary_tra_longer$num_id))){
  subset1 <- binary_tra_longer %>% filter(num_id == j & value>0) %>% pull("aaSeqCDR3")
  vector_overlap <- c()
  for(i in levels(factor(binary_tra_longer$num_id))){
    subset2 <- binary_tra_longer %>% filter(num_id == i & value>0) %>% pull("aaSeqCDR3")
    intersect_rep <- length(intersect(subset1, subset2))
    total <- length(subset2)
    vector_overlap <- c(vector_overlap,intersect_rep/total)
  }
  df <- as.data.frame(x = vector_overlap)
  colnames(df) <- j
  df
  df_all4 <- cbind(df_all4, df)
}

df_all4 <- df_all4[,2:29]
rownames(df_all4) <- colnames(df_all4)

#write.csv(df_all4, "rep_overlap_tra.csv")
```


```{r fig.width=15, fig.height=8}
df24 <- df_all4
df24[df24 == 1] <- 0

### plot dotplot of overlaps

df25 <- df24 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df25, aes(id, factor(name, levels = rev(levels(factor(name)))))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(7,12)) +
  geom_text(aes(label = round(value*100, digits = 1))) + 
  scale_colour_gradient2(low = "lightskyblue", mid = "lightsteelblue2", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) 

```

### Rozdelene na CD4 a CD8

```{r fig.width=7, fig.height=5}

rep_overlap_tra <- read_csv("rep_overlap_tra.csv") %>% column_to_rownames("...1")

rep_overlap_tra_cd4 <- rep_overlap_tra[1:14,1:14]
rep_overlap_tra_cd8 <- rep_overlap_tra[15:28,15:28]

rep_overlap_tra_cd4 <- rep_overlap_tra_cd4[ rev(c(5, 4, 3, 2, 1, 7, 6, 12, 11, 10, 9, 8, 14, 13)),
                                            rev(c(5, 4, 3, 2, 1, 7, 6, 12, 11, 10, 9, 8, 14, 13))]

levels_names <- colnames(rep_overlap_tra_cd4)

df_overlap_tra_cd4 <- rep_overlap_tra_cd4
df_overlap_tra_cd4[df_overlap_tra_cd4 == 1] <- 0

### plot dotplot of overlaps

df_overlap_tra_cd4 <- df_overlap_tra_cd4 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df_overlap_tra_cd4, aes(x = factor(name, levels = levels_names), y = factor(id, levels = rev(levels_names)))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(5,9)) +
  geom_text(aes(label = round(value*100, digits = 1)), size = 3) + 
  scale_colour_gradient2(low = "white", mid = "lightsteelblue", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) + ggtheme()

#ggsave(file = "./final_fig/overlap_cd4_tra.png", width = 20, height = 16, units = "cm")
#ggsave(file = "./final_fig/overlap_cd4_tra.svg", width = 20, height = 16, units = "cm")

```

```{r fig.width=7, fig.height=5}
rep_overlap_tra_cd8 <- rep_overlap_tra_cd8[rev(c(12:14,8:11,5:7,1:4)),rev(c(12:14,8:11,5:7,1:4))]

levels_names <- rev(colnames(rep_overlap_tra_cd8))

df_overlap_tra_cd8 <- rep_overlap_tra_cd8
df_overlap_tra_cd8[df_overlap_tra_cd8 == 1] <- 0

### plot dotplot of overlaps

df_overlap_tra_cd8 <- df_overlap_tra_cd8 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df_overlap_tra_cd8, aes(factor(name, levels = levels_names), factor(id, levels = rev(levels_names)))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(5,9)) +
  geom_text(aes(label = round(value*100, digits = 1)), size = 3) + 
  scale_colour_gradient2(low = "white", mid = "lightsteelblue", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) + ggtheme()

#ggsave(file = "./final_fig/overlap_cd8_tra.png", width = 20, height = 16, units = "cm")
#ggsave(file = "./final_fig/overlap_cd8_tra.svg", width = 20, height = 16, units = "cm")

```

### A la violin

```{r fig.width=3, fig.height=3}
violin_tra_cd4 <- rep_overlap_tra_cd4 %>% 
  rownames_to_column("sample1") %>% 
  pivot_longer(!sample1, names_to = "sample2") %>% 
  mutate(sample2 = stringr::str_replace_all(sample2, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  mutate(sample1 = stringr::str_replace_all(sample1, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  separate(sample1, into = c("Cell_type1","Organ1","Strain1","Exp1"), remove = F)  %>% 
  separate(sample2, into = c("Cell_type2","Organ2","Strain2","Exp2"), remove = F) %>% 
  filter(value<1)


violin_tra_cd8 <- rep_overlap_tra_cd8 %>% 
  rownames_to_column("sample1") %>% 
  pivot_longer(!sample1, names_to = "sample2") %>% 
  mutate(sample2 = stringr::str_replace_all(sample2, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  mutate(sample1 = stringr::str_replace_all(sample1, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  separate(sample1, into = c("Cell_type1","Organ1","Strain1","Exp1"), remove = F)  %>% 
  separate(sample2, into = c("Cell_type2","Organ2","Strain2","Exp2"), remove = F) %>% 
  filter(value<1)

```

```{r fig.width=3, fig.height=3}
violin_tra_cd4 %>% filter(Organ1 == "Thymus" & Organ2 == "Thymus") %>% 
  mutate(organ_strain1 = paste(Organ1, Strain1), 
                          organ_strain2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strain1, organ_strain2, sep = " - ")) %>% 
  filter(comparison %in% c("Thymus WT - Thymus WT", "Thymus CA - Thymus WT", "Thymus CAKR - Thymus WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD4 TRA Thymus") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.35)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd4_tra_thy.png", width = 9.5, height = 9.5,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd4_tra_thy.eps", width = 9.5, height = 9.5,  units = "cm")

```

```{r fig.width=3, fig.height=3}
violin_tra_cd4 %>% filter(Organ1 == "LN" & Organ2 == "LN") %>% 
  mutate(organ_strain1 = paste(Organ1, Strain1), 
                          organ_strain2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strain1, organ_strain2, sep = " - ")) %>% 
  filter(comparison %in% c("LN WT - LN WT", "LN CA - LN WT", "LN CAKR - LN WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD4 TRA LN") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.35)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd4_tra_LN.png", width = 8, height = 8,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd4_tra_LN.eps", width = 8, height = 8,  units = "cm")

```


```{r fig.width=3, fig.height=3}
violin_tra_cd8 %>% filter(Organ1 == "Thymus" & Organ2 == "Thymus") %>% 
  mutate(organ_strain1 = paste(Organ1, Strain1), 
                          organ_strain2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strain1, organ_strain2, sep = " - ")) %>% 
  filter(comparison %in% c("Thymus WT - Thymus WT", "Thymus CA - Thymus WT", "Thymus CAKR - Thymus WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD8 TRA Thymus") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.35)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd8_tra_thy.png", width = 9.5, height = 9.5,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd8_tra_thy.eps", width = 9.5, height = 9.5,  units = "cm")

```

```{r fig.width=3, fig.height=3}
violin_tra_cd8 %>% filter(Organ1 == "LN" & Organ2 == "LN") %>% 
  mutate(organ_strain1 = paste(Organ1, Strain1), 
                          organ_strain2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strain1, organ_strain2, sep = " - ")) %>% 
  filter(comparison %in% c("LN WT - LN WT", "LN CA - LN WT", "LN CAKR - LN WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD8 TRA LN") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.35)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd8_tra_LN.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd8_tra_LN.eps", width = 8, height = 8, units = "cm")

```

### TRB

## Vsechny vzorky se vsema

```{r}

binary_trb <- excel_count_table_trb %>% 
  mutate_at(vars(starts_with("CD")), .funs = binary) 

binary_trb_longer <- binary_trb %>% select(starts_with("CD"), aaSeqCDR3) %>% pivot_longer(!aaSeqCDR3, names_to = "num_id")

df_all4 <- data.frame("")

for(j in levels(factor(binary_trb_longer$num_id))){
  subset1 <- binary_trb_longer %>% filter(num_id == j & value>0) %>% pull("aaSeqCDR3")
  vector_overlap <- c()
  for(i in levels(factor(binary_trb_longer$num_id))){
    subset2 <- binary_trb_longer %>% filter(num_id == i & value>0) %>% pull("aaSeqCDR3")
    intersect_rep <- length(intersect(subset1, subset2))
    total <- length(subset2)
    vector_overlap <- c(vector_overlap,intersect_rep/total)
  }
  df <- as.data.frame(x = vector_overlap)
  colnames(df) <- j
  df
  df_all4 <- cbind(df_all4, df)
}

df_all4 <- df_all4[,2:29]
rownames(df_all4) <- colnames(df_all4)

#write.csv(df_all4, "rep_overlap_trb.csv")
```


```{r fig.width=30, fig.height=25}
df24 <- df_all4
df24[df24 == 1] <- 0

### plot dotplot of overlaps

df25 <- df24 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df25, aes(id, factor(name, levels = rev(levels(factor(name)))))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(7,12)) +
  geom_text(aes(label = round(value*100, digits = 1))) + 
  scale_colour_gradient2(low = "lightskyblue", mid = "lightsteelblue2", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) 


```

### Rozdelene na CD4 a CD8

```{r fig.width=7, fig.height=5}

rep_overlap_trb <- read_csv("rep_overlap_trb.csv") %>% column_to_rownames("...1")

rep_overlap_trb_cd4 <- rep_overlap_trb[1:14,1:14]
rep_overlap_trb_cd8 <- rep_overlap_trb[15:28,15:28]

rep_overlap_trb_cd4 <- rep_overlap_trb_cd4[rev(c(13,14,8:12,6,7,1:5)),rev(c(13,14,8:12,6,7,1:5))]

levels_names <- rev(colnames(rep_overlap_trb_cd4))


df_overlap_trb_cd4 <- rep_overlap_trb_cd4
df_overlap_trb_cd4[df_overlap_trb_cd4 == 1] <- 0

### plot dotplot of overlaps

df_overlap_trb_cd4 <- df_overlap_trb_cd4 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df_overlap_trb_cd4, aes(factor(name, levels = levels_names), factor(id, levels = rev(levels_names)))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(5,9)) +
  geom_text(aes(label = round(value*100, digits = 1)), size = 3) + 
  scale_colour_gradient2(low = "white", mid = "lightsteelblue", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) + ggtheme()

#ggsave(file = "./final_fig/overlap_cd4_trb.png", width = 20, height = 16, units = "cm")
#ggsave(file = "./final_fig/overlap_cd4_trb.svg", width = 20, height = 16, units = "cm")

```

```{r fig.width=7, fig.height=5}
rep_overlap_trb_cd8 <- rep_overlap_trb_cd8[rev(c(12:14,8:11,5:7,1:4)),rev(c(12:14,8:11,5:7,1:4))]

levels_names <- rev(colnames(rep_overlap_trb_cd8))

df_overlap_trb_cd8 <- rep_overlap_trb_cd8
df_overlap_trb_cd8[df_overlap_trb_cd8 == 1] <- 0

### plot dotplot of overlaps

df_overlap_trb_cd8 <- df_overlap_trb_cd8 %>% rownames_to_column("id") %>%  pivot_longer(-id)

g2 <- ggplot(df_overlap_trb_cd8, aes(factor(name, levels = levels_names), factor(id, levels = rev(levels_names)))) + 
  geom_point(aes(size = value*100, colour = value*100)) + 
  theme_bw() 

g2 + scale_size_continuous(range=c(5,9)) +
  geom_text(aes(label = round(value*100, digits = 1)), size = 3) + 
  scale_colour_gradient2(low = "white", mid = "lightsteelblue", high = "salmon") +
  theme(axis.text.x = element_text(angle = 90)) + ggtheme()

#ggsave(file = "./final_fig/overlap_cd8_trb.png", width = 20, height = 16, units = "cm")
#ggsave(file = "./final_fig/overlap_cd8_trb.svg", width = 20, height = 16, units = "cm")

```

### A la violin

```{r}

violin_trb_cd4 <- rep_overlap_trb_cd4 %>% 
  rownames_to_column("sample1") %>% 
  pivot_longer(!sample1, names_to = "sample2") %>% 
  mutate(sample2 = stringr::str_replace_all(sample2, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  mutate(sample1 = stringr::str_replace_all(sample1, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  separate(sample1, into = c("Cell_type1","Organ1","Strain1","Exp1"), remove = F)  %>% 
  separate(sample2, into = c("Cell_type2","Organ2","Strain2","Exp2"), remove = F) %>% 
  filter(value<1)

violin_trb_cd8 <- rep_overlap_trb_cd8 %>% 
  rownames_to_column("sample1") %>% 
  pivot_longer(!sample1, names_to = "sample2") %>% 
  mutate(sample2 = stringr::str_replace_all(sample2, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  mutate(sample1 = stringr::str_replace_all(sample1, pattern = "Lymph nodes", replacement = "LN"))  %>% 
  separate(sample1, into = c("Cell_type1","Organ1","Strain1","Exp1"), remove = F)  %>% 
  separate(sample2, into = c("Cell_type2","Organ2","Strain2","Exp2"), remove = F) %>% 
  filter(value<1)
```

```{r fig.width=5, fig.height=5}
violin_trb_cd8 %>% filter(Organ1 == "Thymus" & Organ2 == "Thymus") %>% 
  mutate(organ_strbin1 = paste(Organ1, Strain1), 
                          organ_strbin2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strbin1, organ_strbin2, sep = " - ")) %>% 
  filter(comparison %in% c("Thymus WT - Thymus WT", "Thymus CA - Thymus WT", "Thymus CAKR - Thymus WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD8 TRB Thymus") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.25)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd8_trb_thy.png", width = 9.5, height = 9.5,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd8_trb_thy.eps", width = 9.5, height = 9.5,  units = "cm")

```

```{r fig.width=4, fig.height=3}
violin_trb_cd8 %>% filter(Organ1 == "LN" & Organ2 == "LN") %>% 
  mutate(organ_strbin1 = paste(Organ1, Strain1), 
                          organ_strbin2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strbin1, organ_strbin2, sep = " - ")) %>% 
  filter(comparison %in% c("LN WT - LN WT", "LN CA - LN WT", "LN CAKR - LN WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD8 TRB LN") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.25)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd8_trb_LN.png", width = 8, height = 8,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd8_trb_LN.eps", width = 8, height = 8,  units = "cm")

```

```{r fig.width=5, fig.height=5}
violin_trb_cd4 %>% filter(Organ1 == "Thymus" & Organ2 == "Thymus") %>% 
  mutate(organ_strbin1 = paste(Organ1, Strain1), 
                          organ_strbin2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strbin1, organ_strbin2, sep = " - ")) %>% 
  filter(comparison %in% c("Thymus WT - Thymus WT", "Thymus CA - Thymus WT", "Thymus CAKR - Thymus WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD4 TRB Thymus") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.20)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd4_trb_thy.png", width = 9.5, height = 9.5,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd4_trb_thy.eps", width = 9.5, height = 9.5,  units = "cm")

```

```{r fig.width=4, fig.height=3}
violin_trb_cd4 %>% filter(Organ1 == "LN" & Organ2 == "LN") %>% 
  mutate(organ_strbin1 = paste(Organ1, Strain1), 
                          organ_strbin2 = paste(Organ2, Strain2)) %>% 
  mutate(comparison = paste(organ_strbin1, organ_strbin2, sep = " - ")) %>% 
  filter(comparison %in% c("LN WT - LN WT", "LN CA - LN WT", "LN CAKR - LN WT")) %>% 
  ggplot(aes(x = comparison, y = value)) +
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0) + 
  geom_jitter(position=position_jitter(0.2), size = 2, aes(color = comparison)) +
  ggtitle("CD4 TRB LN") +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  ylab("% repertoire overlap") + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,0.20)) + xlab("")

#ggsave("./final_fig/overlaps/rep_overlap_cd4_trb_LN.png", width = 8, height = 8,  units = "cm")
#ggsave("./final_fig/overlaps/rep_overlap_cd4_trb_LN.eps", width = 8, height = 8,  units = "cm")

```

# SessionInfo

```{r}
sessionInfo()
```

