---
title: "Analysis of TCR repertoires of LckWT, LckCA and LckCA/KR mice"
author: "Veronika Niederlova"
date: "8/23/2022"
output: rmdformats::material
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE, error = TRUE, cache = TRUE)

library(tidyverse)
library(readr)
library(readxl)
library(cowplot)
library(patchwork)
library(immunarch)
library(pheatmap)
library(factoextra)

ggtheme <- function() {
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10),
    text = element_text(size = 10, colour = "black"),
    legend.text = element_text(size = 10),
    legend.key.size =  unit(10, units = "points")
    
  )
}

```


# Initial processing

This document describes the analysis of TCR repertoires for the manuscript "Unique roles of coreceptor-bound LCK in helper and cytotoxic T cellsby Horkova et al. The analysis workflow included: 

* sorting of CD4 and CD8 cells from lymph nodes and thymi of *Lck^WT^*, *Lck^CA^* and *Lck^CA/KR^* mice,
* isolation of RNA using 
* preparation of TCR-enriched libraries using the [NEBNext® Mouse Immune Sequencing Kit](https://international.neb.com/products/e6330-nebnext-immune-sequencing-kit-mouse#Product%20Information_Reagents-Supplied)
* sequencing the libraries on Illumina Miseq 300bp paired-end
* demultiplexing reads in Illumina BaseSpace
* processing fastq files using the recommended [pRESTO workflow](https://international.neb.com/-/media/nebus/files/nebnext-chart-pdfs/nebnext-immune-sequencing-data-analysis-workflow-mouse.pdf?rev=7f21b23a3ce444e3bc4dcd57282402ba&hash=C0AE101CBA6C31B2A897A207D9EA6A4D) on Galaxy with default parameters
* aligning processed fastq files using [MIXCR](https://mixcr.readthedocs.io/en/develop/)

. This analysis starts with file `merged_TCR_repertoires.csv` which you can [download from Zenodo]( ############# )


```{r}
df <- read.csv("merged_TCR_repertoires.csv")

# Count tables


# Gene segment usage analysis
```

# Analysis of NKT cells

Next, we will identify gene segments that are typical for invariant NKT cells. These segments are TRAV11 and TRAJ18 for TRA and TRBV1, TRBV13-2 and TRBV29 for TRB. Please, note that some of the included TRB chains are used by conventional cells as well. 

Counts of NKT gene segments:

```{r}
# NKT analysis
merged_repertoires <- merged_repertoires %>% 
  mutate(is_nkt = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11")) & (grepl(allJHitsWithScore, pattern = "TRAJ18")) |
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")) |
      (grepl(allVHitsWithScore, pattern = "TRBV1\\*")) |
      (grepl(allVHitsWithScore, pattern = "TRBV29"))  ,"yes","no")) %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% mutate(nkt_trbv13_2 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")),"yes","no")) %>% mutate(nkt_trbv29 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV29")),"yes","no")) %>% mutate(nkt_trbv1 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV1\\*")),"yes","no"))

merged_repertoires %>% group_by(num_id)
```

### All NKT

```{r}

# All NKT
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, is_nkt, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, is_nkt) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(is_nkt == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.eps", width = 8, height = 8, units = "cm")

  
## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRB") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.eps", width = 8, height = 8, units = "cm")


## TRA

# All NKT
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRA") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, is_nkt, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, is_nkt) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(is_nkt == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_all_nkt_tra_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_all_nkt_tra_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRA") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,1)) + xlab("")
  
ggsave("./final_fig/nkt/pct_all_nkt_tra_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_all_nkt_tra_cd8.eps", width = 8, height = 8, units = "cm")

```

### TRAV11 TRAJ18

```{r}
# TRAV11 TRAJ18
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRA") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trav11_traj18, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trav11_traj18) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trav11_traj18 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRA TRAV11 TRAJ18") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

## CD8
nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRA TRAV11 TRAJ18") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,1)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.eps", width = 8, height = 8, units = "cm")
```

### TRBV13-2

```{r}

## TRB

# TRBV13-2
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv13_2, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv13_2) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv13_2 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV13-2") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV13-2") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.eps", width = 8, height = 8, units = "cm")
```

### TRBV29

```{r}
# TRBV29
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv29, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv29) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv29 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV29") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV29") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.eps", width = 8, height = 8, units = "cm")
```

### TRBV1

```{r}

merged_repertoires %>% group_by(nkt_trbv1) %>% tally()

# TRBV1
nkt_table <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv1, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv1) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv1 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

nkt_table

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")

## CD4
nkt_table %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV1") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

nkt_table %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV1") + 
  theme(axis.text.x = element_text(angle = 90)) + 
  theme_classic() + 
  ggtheme() + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.png", width = 8, height = 8, units = "cm")
ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.eps", width = 8, height = 8, units = "cm")

```

```{r}
# PCA 

## 

## 

# Diversity index

# Top20 WT LN sequences

## Summary plot

## Heatmaps

# Repertoire overlap

## 

## 


```

