---
title: "Analysis of TCR repertoires of LckWT, LckCA and LckCA/KR mice"
author: "Veronika Niederlova"
date: "8/23/2022"
output: rmdformats::material
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include = TRUE, warning = FALSE, message = FALSE, error = TRUE, cache = TRUE)

library(tidyverse)
library(readr)
library(readxl)
library(cowplot)
library(patchwork)
library(immunarch)
library(pheatmap)
library(factoextra)
library(kableExtra)

ggtheme <- function() {
  theme(
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 10),
    text = element_text(size = 10, colour = "black"),
    legend.text = element_text(size = 10),
    legend.key.size =  unit(10, units = "points"),
    axis.text.x = element_text(angle = 90)) 
  }

binary <- function(x){
  x <- ifelse(x>0,1,0)
  return(x)
}

```


# Initial processing

This document describes the analysis of TCR repertoires for the manuscript "Unique roles of coreceptor-bound LCK in helper and cytotoxic T cellsby Horkova et al. The analysis workflow included: 

* sorting of CD4 and CD8 cells from lymph nodes and thymi of *Lck^WT^*, *Lck^CA^* and *Lck^CA/KR^* mice,
* isolation of RNA using 
* preparation of TCR-enriched libraries using the [NEBNext® Mouse Immune Sequencing Kit](https://international.neb.com/products/e6330-nebnext-immune-sequencing-kit-mouse#Product%20Information_Reagents-Supplied)
* sequencing the libraries on Illumina Miseq 300bp paired-end
* demultiplexing reads in Illumina BaseSpace
* processing fastq files using the recommended [pRESTO workflow](https://international.neb.com/-/media/nebus/files/nebnext-chart-pdfs/nebnext-immune-sequencing-data-analysis-workflow-mouse.pdf?rev=7f21b23a3ce444e3bc4dcd57282402ba&hash=C0AE101CBA6C31B2A897A207D9EA6A4D) on Galaxy with default parameters
* aligning processed fastq files using [MIXCR](https://mixcr.readthedocs.io/en/develop/)

Raw data are deposited in the Sequence Read Archive (PRJNA872031). Processed files can be downloaded [download from Zenodo]( ############# ).
. This analysis starts with file `merged_TCR_repertoires.csv` which you can [download from Zenodo]( ############# )


```{r}
merged_repertoires <- read.csv("merged_TCR_repertoires.csv")
md <- read.csv("metadata_Lck.csv")

head(md)
```

# Count tables

```{r}
counts_tra <- read_csv("count_table_tra2.csv")
counts_trb <- read_csv("count_table_trb2.csv")

```

## Prop table without NKT without less than 5 counts


# count table

## TRB

Nice table for excel

```{r}
md2_trb <- merged_repertoires %>% filter(chain == "TRB") %>% group_by(aaSeqCDR3) %>% slice_head(n = 1)

excel_count_table_trb <- counts_trb %>% left_join(md2_trb %>% select(allVHitsWithScore, allDHitsWithScore, allJHitsWithScore))

md_to_join <- md %>% mutate(new_name = paste(Cell_type, Organ, Mouse_strain, Exp))  %>% select(num_id, new_name) %>% arrange(num_id)
  
colnames(excel_count_table_trb)[2:29] <- c(
  "CD8 Lymph nodes WT Exp01",
"CD8 Thymus WT Exp01",
"CD4 Thymus CA Exp03",
"CD8 Lymph nodes WT Exp03",
"CD4 Lymph nodes CAKR Exp03",
"CD8 Lymph nodes CA Exp03",
"CD4 Thymus WT Exp02",
"CD8 Lymph nodes CA Exp02",
"CD8 Thymus CA Exp03",
"CD4 Lymph nodes CAKR Exp02",
"CD4 Lymph nodes CA Exp02",
"CD4 Lymph nodes WT Exp03",
"CD4 Thymus CAKR Exp03",
"CD8 Thymus WT Exp02",
"CD8 Lymph nodes CAKR Exp02",
"CD4 Thymus CAKR Exp02",
"CD4 Thymus CA Exp01",
"CD8 Thymus CAKR Exp03",
"CD8 Thymus CAKR Exp02",
"CD8 Lymph nodes WT Exp02",
"CD4 Lymph nodes CA Exp01",
"CD4 Thymus WT Exp03",
"CD8 Lymph nodes CAKR Exp03",
"CD4 Lymph nodes CA Exp03",
"CD4 Lymph nodes WT Exp02",
"CD8 Thymus WT Exp03",
"CD4 Thymus CAKR Exp01",
"CD8 Thymus CA Exp01")

excel_count_table_trb2 <- excel_count_table_trb %>% dplyr::select(1, 31, 32, 30, 33, 22, 12, 25, 11, 6, 26, 13, 18, 4, 28, 17, 14, 8, 23, 9, 7, 16, 24, 2, 21, 5, 29, 10, 20, 19, 3, 15, 27)

excel_count_table_trb3 <- excel_count_table_trb2 %>% mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no"))

count_table_trb4 <- as.matrix(excel_count_table_trb3[,6:33])
rownames(count_table_trb4) <- excel_count_table_trb3$aaSeqCDR3

trb4_norm <- scale(count_table_trb4, center=FALSE, scale=colSums(count_table_trb4))

prop.table.trb <- cbind(trb4_norm, excel_count_table_trb %>% select(-starts_with("CD")) ) 

prop.table.trb2 <- prop.table.trb %>% dplyr::select(29, 31, 32, 30, 1:28)

```

## TRA

```{r}
md2_tra <- merged_repertoires %>% filter(chain == "TRA") %>% group_by(aaSeqCDR3) %>% slice_head(n = 1)

excel_count_table_tra <- counts_tra %>% left_join(md2_tra %>% select(allVHitsWithScore, allDHitsWithScore, allJHitsWithScore))

md_to_join <- md %>% mutate(new_name = paste(Cell_type, Organ, Mouse_strain, Exp))  %>% select(num_id, new_name) %>% arrange(num_id)
  
colnames(excel_count_table_tra)[2:29] <- c(
  "CD8 Lymph nodes WT Exp01",
"CD8 Thymus WT Exp01",
"CD4 Thymus CA Exp03",
"CD8 Lymph nodes WT Exp03",
"CD4 Lymph nodes CAKR Exp03",
"CD8 Lymph nodes CA Exp03",
"CD4 Thymus WT Exp02",
"CD8 Lymph nodes CA Exp02",
"CD8 Thymus CA Exp03",
"CD4 Lymph nodes CAKR Exp02",
"CD4 Lymph nodes CA Exp02",
"CD4 Lymph nodes WT Exp03",
"CD4 Thymus CAKR Exp03",
"CD8 Thymus WT Exp02",
"CD8 Lymph nodes CAKR Exp02",
"CD4 Thymus CAKR Exp02",
"CD4 Thymus CA Exp01",
"CD8 Thymus CAKR Exp03",
"CD8 Thymus CAKR Exp02",
"CD8 Lymph nodes WT Exp02",
"CD4 Lymph nodes CA Exp01",
"CD4 Thymus WT Exp03",
"CD8 Lymph nodes CAKR Exp03",
"CD4 Lymph nodes CA Exp03",
"CD4 Lymph nodes WT Exp02",
"CD8 Thymus WT Exp03",
"CD4 Thymus CAKR Exp01",
"CD8 Thymus CA Exp01")

excel_count_table_tra2 <- excel_count_table_tra %>% dplyr::select(1, 31, 32, 30, 33, 22, 12, 25, 11, 6, 26, 13, 18, 4, 28, 17, 14, 8, 23, 9, 7, 16, 24, 2, 21, 5, 29, 10, 20, 19, 3, 15, 27)

excel_count_table_tra3 <- excel_count_table_tra2 %>% mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no"))

excel_count_table_tra4 <- excel_count_table_tra3 %>% filter(nkt_trav11_traj18 == "no")

count_table_tra4 <- as.matrix(excel_count_table_tra4[,6:33])
rownames(count_table_tra4) <- excel_count_table_tra4$aaSeqCDR3

tra4_norm <- scale(count_table_tra4, center=FALSE, scale=colSums(count_table_tra4))

prop.table.tra <- cbind(tra4_norm, excel_count_table_tra4 %>% select(-starts_with("CD")) ) 

prop.table.tra2 <- prop.table.tra %>% dplyr::select(29, 31, 32, 30, 1:28)

```

# less than 5 removed

## CD4

Remove CDR3s with occurrence in less than 5 samples

```{r}

# TRA
keep_tra_cd4 <- excel_count_table_tra %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") | 
    (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% 
  filter(nkt_trav11_traj18 == "no") %>% 
  select(aaSeqCDR3, starts_with("CD4")) %>% 
  mutate_at(vars(starts_with("CD4")), .funs = binary) 
 
keep_tra_cd4$sum <- rowSums((keep_tra_cd4 %>% select(-aaSeqCDR3)))

keep_tra_cd4$keep <- ifelse(keep_tra_cd4$sum>4,1,0)

keep_tra_cd4_sequences <- pull(keep_tra_cd4 %>% filter(keep == 1), aaSeqCDR3)

# TRB
keep_trb_cd4 <- excel_count_table_trb %>% select(aaSeqCDR3, starts_with("CD4")) %>%  mutate_at(vars(starts_with("CD4")), .funs = binary) 
 
keep_trb_cd4$sum <- rowSums((keep_trb_cd4 %>% select(-aaSeqCDR3)))

keep_trb_cd4$keep <- ifelse(keep_trb_cd4$sum>4,1,0)

keep_trb_cd4_sequences <- pull(keep_trb_cd4 %>% filter(keep == 1), aaSeqCDR3)

```

## CD8

Remove CDR3s with occurrence in less than 5 samples

```{r}

# TRA
keep_tra_cd8 <- excel_count_table_tra %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") | 
    (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% 
  filter(nkt_trav11_traj18 == "no") %>% 
  select(aaSeqCDR3, starts_with("CD8")) %>% 
  mutate_at(vars(starts_with("CD8")), .funs = binary) 

keep_tra_cd8$sum <- rowSums((keep_tra_cd8 %>% select(-aaSeqCDR3)))

keep_tra_cd8$keep <- ifelse(keep_tra_cd8$sum>4,1,0)

keep_tra_cd8_sequences <- pull(keep_tra_cd8 %>% filter(keep == 1), aaSeqCDR3)

# TRB
keep_trb_cd8 <- excel_count_table_trb %>% select(aaSeqCDR3, starts_with("CD8")) %>%  mutate_at(vars(starts_with("CD8")), .funs = binary) 
 
keep_trb_cd8$sum <- rowSums((keep_trb_cd8 %>% select(-aaSeqCDR3)))

keep_trb_cd8$keep <- ifelse(keep_trb_cd8$sum>4,1,0)

keep_trb_cd8_sequences <- pull(keep_trb_cd8 %>% filter(keep == 1), aaSeqCDR3)

```

## TRB

```{r}

excel_count_table_trb3_filter <- excel_count_table_trb2  %>% filter(aaSeqCDR3 %in% keep_trb_cd4_sequences | aaSeqCDR3 %in% keep_trb_cd8_sequences) 

count_table_trb4_filter <- as.matrix(excel_count_table_trb3_filter[,6:33])
rownames(count_table_trb4_filter) <- excel_count_table_trb3_filter$aaSeqCDR3

trb4_norm_filter <- scale(count_table_trb4_filter, center=FALSE, scale=colSums(count_table_trb4_filter))

prop.table.trb_filter <- cbind(trb4_norm_filter, excel_count_table_trb3_filter %>% select(-starts_with("CD")) ) 

prop.table.trb2_filter <- prop.table.trb_filter %>% dplyr::select(29, 31, 32, 30, 1:28)
```

## TRA

```{r}
excel_count_table_tra3_filter <- excel_count_table_tra2  %>% filter(aaSeqCDR3 %in% keep_tra_cd4_sequences | aaSeqCDR3 %in% keep_tra_cd8_sequences) %>% filter(is_nkt == "no")

count_table_tra4_filter <- as.matrix(excel_count_table_tra3_filter[,6:33])
rownames(count_table_tra4_filter) <- excel_count_table_tra3_filter$aaSeqCDR3

tra4_norm_filter <- scale(count_table_tra4_filter, center=FALSE, scale=colSums(count_table_tra4_filter))

prop.table.tra_filter <- cbind(tra4_norm_filter, excel_count_table_tra3_filter %>% select(-starts_with("CD")) ) 


prop.table.tra2_filter <- prop.table.tra_filter 
```

# Gene segment usage analysis

```{r}


```

# Analysis of NKT cells

Next, we will identify gene segments that are typical for invariant NKT cells. These segments are TRAV11 and TRAJ18 for TRA and TRBV1, TRBV13-2 and TRBV29 for TRB. Please, note that some of the included TRB chains are used by conventional cells as well. 

Counts of NKT gene segments:

```{r}
# NKT analysis
merged_repertoires <- merged_repertoires %>% 
  mutate(is_nkt = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11")) & (grepl(allJHitsWithScore, pattern = "TRAJ18")) |
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")) |
      (grepl(allVHitsWithScore, pattern = "TRBV1\\*")) |
      (grepl(allVHitsWithScore, pattern = "TRBV29"))  ,"yes","no")) %>% 
  mutate(nkt_trav11_traj18 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRAV11") & (grepl(allJHitsWithScore, pattern = "TRAJ18"))),"yes","no")) %>% mutate(nkt_trbv13_2 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV13-2")),"yes","no")) %>% mutate(nkt_trbv29 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV29")),"yes","no")) %>% mutate(nkt_trbv1 = if_else(
    (grepl(allVHitsWithScore, pattern = "TRBV1\\*")),"yes","no"))

md <- md %>% select(Exp, Organ, Cell_type, Mouse_strain, num_id) %>% mutate_at(vars("num_id"), as.numeric)

# All TRB NKT
nkt_table_trb <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, is_nkt, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, is_nkt) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(is_nkt == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRAV11 TRAJ18
nkt_table_trav11_traj18 <- merged_repertoires %>% 
   filter(chain == "TRA") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trav11_traj18, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trav11_traj18) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trav11_traj18 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRBV1
nkt_table_trbv1 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv1, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv1) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv1 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

# TRBV13-2
nkt_table_trbv13_2 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv13_2, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv13_2) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv13_2 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 


# TRBV29
nkt_table_trbv29 <- merged_repertoires %>% 
   filter(chain == "TRB") %>% 
 mutate(new_name = paste(Exp, Organ, Cell_type, Mouse_strain)) %>% 
  dplyr::select(new_name, cloneCount, nkt_trbv29, num_id) %>% 
  uncount(cloneCount) %>% 
  group_by(num_id, nkt_trbv29) %>% 
    summarise(n = n()) %>% 
mutate(freq = n / sum(n)) %>% dplyr::filter(nkt_trbv29 == "yes") %>% 
  ungroup %>% 
  left_join(md) %>% 
  unique %>%
   mutate(sample_type = paste(Cell_type, Organ, Mouse_strain)) 

levels_cd4 <- c("CD4 Thymus WT", "CD4 Thymus CA", "CD4 Thymus CAKR", "CD4 Lymph nodes WT", "CD4 Lymph nodes CA", "CD4 Lymph nodes CAKR")
levels_cd8 <- c("CD8 Thymus WT", "CD8 Thymus CA", "CD8 Thymus CAKR", "CD8 Lymph nodes WT", "CD8 Lymph nodes CA", "CD8 Lymph nodes CAKR")

```


```{r}
nkt_table_trb$nkt_sample <- "all_nkt_trb"
nkt_table_trav11_traj18$nkt_sample <- "trav11_traj18"
nkt_table_trbv1$nkt_sample <- "trbv1"
nkt_table_trbv13_2$nkt_sample <- "trbv13_2"
nkt_table_trbv29$nkt_sample <- "trbv29"

nkt_table <- cbind(nkt_table_trb$num_id, nkt_table_trb$sample_type, nkt_table_trb$Exp, nkt_table_trb$freq, 
                   nkt_table_trav11_traj18$freq, nkt_table_trbv1$freq, nkt_table_trbv13_2$freq,
                   nkt_table_trbv29$freq) %>% as.data.frame

colnames(nkt_table) <- c("num_id","Sample","Exp","All TRB NKT", "TRAV11-TRAJ18", "TRBV1","TRBV13-2","TRBV29")
nkt_table <- nkt_table %>% mutate(Sample = paste(Sample, Exp)) %>% select(-num_id, -Exp) %>% arrange(Sample)

kable(nkt_table, format = "html") %>%
  kable_styling(full_width = F, font_size = 11, 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

### All NKT

```{r fig.width=7, fig.height=3.5}
## CD4
p01 <- nkt_table_trb %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("NKT segments in CD4 TRB") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p02 <- nkt_table_trb %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("NKT segments in CD8 TRB") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_all_nkt_trb_cd8.eps", width = 8, height = 8, units = "cm")

p01 + p02
```


### TRAV11 TRAJ18

```{r fig.width=7, fig.height=3.5}
## CD4
p03 <- nkt_table_trav11_traj18 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRAV11-TRAJ18 segments in CD4 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p04 <- nkt_table_trav11_traj18 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRAV11-TRAJ18 segments in CD8 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,1)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trav11_traj18_tra_cd8.eps", width = 8, height = 8, units = "cm")

p03 + p04
```

### TRBV1

```{r fig.width=7, fig.height=3.5}

## CD4
p05 <- nkt_table_trbv1 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV1") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p05 <- nkt_table_trbv1 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV1") + 
  theme_classic() + 
  ggtheme() + 
  
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv1_cd8.eps", width = 8, height = 8, units = "cm")

p05 + p06
```

### TRBV13-2

```{r fig.width=7, fig.height=3.5}

## CD4
p07 <- nkt_table_trbv13_2 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRBV13-2 segments in CD4 cells") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p08 <- nkt_table_trbv13_2 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("TRBV13-2 segments in CD8 cells") + 
  theme_classic() + 
  ggtheme() + 
  
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv13_2_cd8.eps", width = 8, height = 8, units = "cm")

p07 + p08
```

### TRBV29

```{r fig.width=7, fig.height=3.5}

## CD4
p09 <- nkt_table_trbv29 %>% 
  filter(Cell_type == "CD4") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd4))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD4 TRBV29") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd4.eps", width = 8, height = 8, units = "cm")

## CD8
p10 <- nkt_table_trbv29 %>% 
  filter(Cell_type == "CD8") %>% 
  ggplot(aes(y = freq*100, x = factor(sample_type, levels = levels_cd8))) +
  stat_summary(fun = "mean",
               geom = "crossbar", 
               width = 0.5,
               colour = "black") +
  geom_point(aes(color = Mouse_strain)) + 
  ylab("Percentage of all chains") + 
  ggtitle("CD8 TRBV29") + 
  theme_classic() + 
  ggtheme() + 
  scale_color_manual(values = c("dodgerblue","indianred2","gray40")) + 
  ylim(c(0,NA)) + xlab("")
  
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.png", width = 8, height = 8, units = "cm")
#ggsave("./final_fig/nkt/pct_nkt_trbv29_cd8.eps", width = 8, height = 8, units = "cm")

p09 + p10
```

# PCA 




## TRA + TRB together

CD4

```{r}
colnames(prop.table.tra2_filter.cd4) == colnames(prop.table.trb2_filter.cd4)
prop.table.tra2_filter.merge <- rbind(prop.table.tra2_filter.cd4, prop.table.trb2_filter.cd4)
  
res.pca.merge.cd4 <- prcomp(t(prop.table.tra2_filter.merge), scale = TRUE, center = T)
fviz_eig(res.pca.merge.cd4)

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.merge.cd4  <-  colnames(prop.table.tra2_filter.merge[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Organ), # color by groups
             legend.title = "Groups",
             repel = TRUE, invisible="quali")

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE, invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","black")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd4$Organ),
                  color = as.factor(mdres.pca.merge.cd4$Strain)))

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd4$Organ),
                  color = as.factor(mdres.pca.merge.cd4$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  ggtheme()
#ggsave("final_fig/pca/cd4.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4.svg", width = 2.6, height = 1.7)

fviz_pca_ind(res.pca.merge.cd4,
             col.ind = as.factor(mdres.pca.merge.cd4$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd4$Organ),
                  color = as.factor(mdres.pca.merge.cd4$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  xlim(c(-30,0)) + ylim(-2,15) +
  ggtheme()


#ggsave("final_fig/pca/cd4_zoom.png", width = 2.2, height = 1.4)
#ggsave("final_fig/pca/cd4_zoom.svg", width = 2.2, height = 1.4)
```


CD8

```{r}
colnames(prop.table.tra2_filter.cd8) == colnames(prop.table.trb2_filter.cd8)
prop.table.tra2_filter.merge <- rbind(prop.table.tra2_filter.cd8, prop.table.trb2_filter.cd8)
  
res.pca.merge.cd8 <- prcomp(t(prop.table.tra2_filter.merge), scale = TRUE, center = T)
fviz_eig(res.pca.merge.cd8)

fviz_pca_ind(res.pca.merge.cd8,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.merge.cd8  <-  colnames(prop.table.tra2_filter.merge[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.merge.cd8,
             col.ind = as.factor(mdres.pca.merge.cd8$Strain), # color by groups
             legend.title = "Groups",
             invisible="quali", pointsize = 0,
             geom = "point") +
   scale_color_manual(values=c("dodgerblue","indianred2","black")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd8$Organ),
                  color = as.factor(mdres.pca.merge.cd8$Strain))) + 
  scale_shape_manual(values = c(1,1,1,15,15)) +
  ggtheme()

fviz_pca_ind(res.pca.merge.cd8,
             col.ind = as.factor(mdres.pca.merge.cd8$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd8$Organ),
                  color = as.factor(mdres.pca.merge.cd8$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  ggtheme()
#ggsave("final_fig/pca/cd8.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8.svg", width = 2.6, height = 1.7)

fviz_pca_ind(res.pca.merge.cd8,
             col.ind = as.factor(mdres.pca.merge.cd8$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.merge.cd8$Organ),
                  color = as.factor(mdres.pca.merge.cd8$Strain))) + 
  scale_shape_manual(values = c(8,8,8,15,15)) +
  xlim(c(-32,-18)) + ylim(-4,5) +
  ggtheme()

#ggsave("final_fig/pca/cd8_zoom.png", width = 2.2, height = 1.4)
#ggsave("final_fig/pca/cd8_zoom.svg", width = 2.2, height = 1.4)
```

## TRA

```{r}
res.pca <- prcomp(t(prop.table.tra2_filter[,1:28]), scale = TRUE)

fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


md  <-  colnames(prop.table.tra2_filter[,1:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
md
fviz_pca_ind(res.pca,
             col.ind = as.factor(md$Cell_type), # color by groups
             legend.title = "Groups",
             repel = TRUE)
```

CD4
```{r}
## All

prop.table.tra2_filter.cd4 <- prop.table.tra2_filter[,1:14]
prop.table.tra2_filter.cd4$sum <- rowSums(prop.table.tra2_filter.cd4)

prop.table.tra2_filter.cd4 <- prop.table.tra2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4 <- prcomp(t(prop.table.tra2_filter.cd4), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd4)

fviz_pca_ind(res.pca.tra.cd4,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd4  <-  colnames(prop.table.tra2_filter[,1:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd4,
             col.ind = as.factor(mdres.pca.tra.cd4$Organ), # color by groups
             legend.title = "Groups",
             repel = TRUE)

# LN
prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd4.ln <- prop.table.tra2_filter[,1:7]
prop.table.tra2_filter.cd4.ln$sum <- rowSums(prop.table.tra2_filter.cd4.ln)

prop.table.tra2_filter.cd4.ln <- prop.table.tra2_filter.cd4.ln %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4.ln <- prcomp(t(prop.table.tra2_filter.cd4.ln), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd4.ln)

fviz_pca_ind(res.pca.tra.cd4.ln,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd4.ln  <-  colnames(prop.table.tra2_filter[,1:7])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd4.ln,
             col.ind = as.factor(mdres.pca.tra.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.tra.cd4.ln,
             col.ind = as.factor(mdres.pca.tra.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd4.ln$Organ),
                  color = as.factor(mdres.pca.tra.cd4.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_tra_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_tra_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.tra2_filter.cd4.thy <- prop.table.tra2_filter[,8:14]
prop.table.tra2_filter.cd4.thy$sum <- rowSums(prop.table.tra2_filter.cd4.thy)

prop.table.tra2_filter.cd4.thy <- prop.table.tra2_filter.cd4.thy %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd4.thy <- prcomp(t(prop.table.tra2_filter.cd4.thy), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd4.thy)

fviz_pca_ind(res.pca.tra.cd4.thy,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd4.thy  <-  colnames(prop.table.tra2_filter[,8:14])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd4.thy,
             col.ind = as.factor(mdres.pca.tra.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.tra.cd4.thy,
             col.ind = as.factor(mdres.pca.tra.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd4.thy$Organ),
                  color = as.factor(mdres.pca.tra.cd4.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") +
  ggtheme()
#ggsave("final_fig/pca/cd4_tra_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_tra_thy.svg", width = 2.6, height = 1.7)

```

CD8
```{r}
## All

prop.table.tra2_filter.cd8 <- prop.table.tra2_filter[,15:28]
prop.table.tra2_filter.cd8$sum <- rowSums(prop.table.tra2_filter.cd8)

prop.table.tra2_filter.cd8 <- prop.table.tra2_filter.cd8 %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8 <- prcomp(t(prop.table.tra2_filter.cd8), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd8)

fviz_pca_ind(res.pca.tra.cd8,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd8  <-  colnames(prop.table.tra2_filter[,15:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd8,
             col.ind = as.factor(mdres.pca.tra.cd8$Organ), # color by groups
             legend.title = "Groups",
             repel = TRUE)


# LN
prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd8.ln <- prop.table.tra2_filter[,15:21]
prop.table.tra2_filter.cd8.ln$sum <- rowSums(prop.table.tra2_filter.cd8.ln)

prop.table.tra2_filter.cd8.ln <- prop.table.tra2_filter.cd8.ln %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8.ln <- prcomp(t(prop.table.tra2_filter.cd8.ln), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd8.ln)

fviz_pca_ind(res.pca.tra.cd8.ln,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd8.ln  <-  colnames(prop.table.tra2_filter[,15:21])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd8.ln,
             col.ind = as.factor(mdres.pca.tra.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.tra.cd8.ln,
             col.ind = as.factor(mdres.pca.tra.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd8.ln$Organ),
                  color = as.factor(mdres.pca.tra.cd8.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_tra_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_tra_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.tra2_filter %>% colnames
prop.table.tra2_filter.cd8.thy <- prop.table.tra2_filter[,22:28]
prop.table.tra2_filter.cd8.thy$sum <- rowSums(prop.table.tra2_filter.cd8.thy)

prop.table.tra2_filter.cd8.thy <- prop.table.tra2_filter.cd8.thy %>% filter(sum >0) %>% select(-sum)

res.pca.tra.cd8.thy <- prcomp(t(prop.table.tra2_filter.cd8.thy), scale = TRUE, center = T)
fviz_eig(res.pca.tra.cd8.thy)

fviz_pca_ind(res.pca.tra.cd8.thy,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.tra.cd8.thy  <-  colnames(prop.table.tra2_filter[,22:28])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.tra.cd8.thy,
             col.ind = as.factor(mdres.pca.tra.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.tra.cd8.thy,
             col.ind = as.factor(mdres.pca.tra.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.tra.cd8.thy$Organ),
                  color = as.factor(mdres.pca.tra.cd8.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_tra_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_tra_thy.svg", width = 2.6, height = 1.7)

```

## TRB

```{r}
res.pca.trb <- prcomp(t(prop.table.trb2_filter[,5:32]), scale = TRUE)

fviz_pca_ind(res.pca.trb,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


md_trb  <-  colnames(prop.table.trb2_filter[,5:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  
fviz_pca_ind(res.pca.trb,
             col.ind = as.factor(md_trb$Cell_type), # color by groups
             legend.title = "Groups",
             repel = TRUE)
```

CD4
```{r}
## All
colnames(prop.table.trb2_filter)

prop.table.trb2_filter.cd4 <- prop.table.trb2_filter[,5:18]
prop.table.trb2_filter.cd4$sum <- rowSums(prop.table.trb2_filter.cd4)

prop.table.trb2_filter.cd4 <- prop.table.trb2_filter.cd4 %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4 <- prcomp(t(prop.table.trb2_filter.cd4), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd4)

fviz_pca_ind(res.pca.trb.cd4,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd4  <-  colnames(prop.table.trb2_filter[,5:18])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd4,
             col.ind = as.factor(mdres.pca.trb.cd4$Organ), # color by groups
             legend.title = "Groups",
             repel = TRUE)

# LN
prop.table.trb2_filter %>% colnames
prop.table.trb2_filter.cd4.ln <- prop.table.trb2_filter[,5:11]
prop.table.trb2_filter.cd4.ln$sum <- rowSums(prop.table.trb2_filter.cd4.ln)

prop.table.trb2_filter.cd4.ln <- prop.table.trb2_filter.cd4.ln %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4.ln <- prcomp(t(prop.table.trb2_filter.cd4.ln), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd4.ln)

fviz_pca_ind(res.pca.trb.cd4.ln,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd4.ln  <-  colnames(prop.table.trb2_filter[,5:11])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd4.ln,
             col.ind = as.factor(mdres.pca.trb.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.trb.cd4.ln,
             col.ind = as.factor(mdres.pca.trb.cd4.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd4.ln$Organ),
                  color = as.factor(mdres.pca.trb.cd4.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_trb_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_trb_ln.svg", width = 2.6, height = 1.7)

# Thymus

prop.table.trb2_filter %>% colnames
prop.table.trb2_filter.cd4.thy <- prop.table.trb2_filter[,12:18]
prop.table.trb2_filter.cd4.thy$sum <- rowSums(prop.table.trb2_filter.cd4.thy)

prop.table.trb2_filter.cd4.thy <- prop.table.trb2_filter.cd4.thy %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd4.thy <- prcomp(t(prop.table.trb2_filter.cd4.thy), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd4.thy)

fviz_pca_ind(res.pca.trb.cd4.thy,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd4.thy  <-  colnames(prop.table.trb2_filter[,12:18])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd4.thy,
             col.ind = as.factor(mdres.pca.trb.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.trb.cd4.thy,
             col.ind = as.factor(mdres.pca.trb.cd4.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd4.thy$Organ),
                  color = as.factor(mdres.pca.trb.cd4.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd4_trb_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd4_trb_thy.svg", width = 2.6, height = 1.7)

```
CD8
```{r}
## All

prop.table.trb2_filter.cd8 <- prop.table.trb2_filter[,19:32]
prop.table.trb2_filter.cd8$sum <- rowSums(prop.table.trb2_filter.cd8)

prop.table.trb2_filter.cd8 <- prop.table.trb2_filter.cd8 %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8 <- prcomp(t(prop.table.trb2_filter.cd8), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd8)

fviz_pca_ind(res.pca.trb.cd8,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd8  <-  colnames(prop.table.trb2_filter[,19:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd8,
             col.ind = as.factor(mdres.pca.trb.cd8$Organ), # color by groups
             legend.title = "Groups",
             repel = TRUE)



# LN
prop.table.trb2_filter %>% colnames
prop.table.trb2_filter.cd8.ln <- prop.table.trb2_filter[,19:25]
prop.table.trb2_filter.cd8.ln$sum <- rowSums(prop.table.trb2_filter.cd8.ln)

prop.table.trb2_filter.cd8.ln <- prop.table.trb2_filter.cd8.ln %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8.ln <- prcomp(t(prop.table.trb2_filter.cd8.ln), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd8.ln)

fviz_pca_ind(res.pca.trb.cd8.ln,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd8.ln  <-  colnames(prop.table.trb2_filter[,19:25])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd8.ln,
             col.ind = as.factor(mdres.pca.trb.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.trb.cd8.ln,
             col.ind = as.factor(mdres.pca.trb.cd8.ln$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd8.ln$Organ),
                  color = as.factor(mdres.pca.trb.cd8.ln$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_trb_ln.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_trb_ln.svg", width = 2.6, height = 1.7)


# Thymus

prop.table.trb2_filter %>% colnames
prop.table.trb2_filter.cd8.thy <- prop.table.trb2_filter[,26:32]
prop.table.trb2_filter.cd8.thy$sum <- rowSums(prop.table.trb2_filter.cd8.thy)

prop.table.trb2_filter.cd8.thy <- prop.table.trb2_filter.cd8.thy %>% filter(sum >0) %>% select(-sum)

res.pca.trb.cd8.thy <- prcomp(t(prop.table.trb2_filter.cd8.thy), scale = TRUE, center = T)
fviz_eig(res.pca.trb.cd8.thy)

fviz_pca_ind(res.pca.trb.cd8.thy,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )


mdres.pca.trb.cd8.thy  <-  colnames(prop.table.trb2_filter[,26:32])  %>% 
  as.data.frame() %>% 
  mutate(sample = stringr::str_replace_all(., pattern = "Lymph nodes", replacement = "LN")) %>% 
  separate(sample, into = c("Cell_type","Organ","Strain","Exp"))
  

fviz_pca_ind(res.pca.trb.cd8.thy,
             col.ind = as.factor(mdres.pca.trb.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             repel = TRUE)

fviz_pca_ind(res.pca.trb.cd8.thy,
             col.ind = as.factor(mdres.pca.trb.cd8.thy$Strain), # color by groups
             legend.title = "Groups",
             geom.ind = "point",
             invisible="quali", pointsize = 0) +
   scale_color_manual(values=c("dodgerblue","indianred2","gray40")) +
   geom_point(aes(shape = as.factor(mdres.pca.trb.cd8.thy$Organ),
                  color = as.factor(mdres.pca.trb.cd8.thy$Strain))) +
  geom_hline(yintercept = 0, color = "grey20") +
  geom_vline(xintercept = 0, color = "grey20") 
#ggsave("final_fig/pca/cd8_trb_thy.png", width = 2.6, height = 1.7)
#ggsave("final_fig/pca/cd8_trb_thy.svg", width = 2.6, height = 1.7)

```

# Diversity index

# Top20 WT LN sequences

## Summary plot

## Heatmaps

## Repertoire overlap



```{r}



```

# SessionInfo

```{r}
sessionInfo()
```

